---
title: "Figure4"
author: "Tülay Karakulak"
date: "2024-03-15"
output: html_document
---


```{r libraries}
library(dplyr)
library(ggplot2)
library(knitr)
library(rmarkdown)
library(harmony)
library(Matrix)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(ggpubr)
theme_set(theme_cowplot())
set.seed(12345)
```


```{r ReadIds}
IsoMatch_All <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/TamaIDs/PbIds_TamaIds.tsv', sep=' ', header=FALSE)
head(IsoMatch_All)
```


```{r Unnest}
library(tidyr)
colnames(IsoMatch_All) <- c('TamaID', 'Sample', 'PBid')
head(IsoMatch_All)
```


```{r mergeAnnotations}
isoform_annotations_normal <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/Normal/scisoseq_classification.filtered_lite_classification.txt', header = TRUE, sep='\t')
# ccRCC_2
isoform_annotations_ccRCC_2 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC2/scisoseq_classification.filtered_lite_classification.txt', header = TRUE, sep='\t')
# ccRCC_3
isoform_annotations_ccRCC_3 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC3/scisoseq_classification.filtered_lite_classification.txt', header = TRUE, sep='\t')

isoform_annotations_ccRCC_4 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC4/scisoseq_classification.filtered_lite_classification.txt', header = TRUE, sep='\t')

isoform_annotations_ccRCC_5 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC5/scisoseq_classification.filtered_lite_classification.txt', header = TRUE, sep='\t')

isoform_annotations_normal_tama <- merge(isoform_annotations_normal, IsoMatch_All[IsoMatch_All$Sample == 'Normal',], by.x = 'isoform', by.y='PBid')
isoform_annotations_ccRCC2_tama <- merge(isoform_annotations_ccRCC_2, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC2',], by.x = 'isoform', by.y='PBid')
isoform_annotations_ccRCC3_tama <- merge(isoform_annotations_ccRCC_3, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC3',], by.x = 'isoform', by.y='PBid')
isoform_annotations_ccRCC4_tama <- merge(isoform_annotations_ccRCC_4, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC4',], by.x = 'isoform', by.y='PBid')
isoform_annotations_ccRCC5_tama <- merge(isoform_annotations_ccRCC_5, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC5',], by.x = 'isoform', by.y='PBid')
```


## Check the transcripts found in more than 3 cells
```{r countCells}
isoform_annotations_normal_tama$number_of_Cellbarcodes <- sapply(strsplit(isoform_annotations_normal_tama$cell_barcodes, ","), length)
isoform_annotations_ccRCC2_tama$number_of_Cellbarcodes <- sapply(strsplit(isoform_annotations_ccRCC2_tama$cell_barcodes, ","), length)
isoform_annotations_ccRCC3_tama$number_of_Cellbarcodes <- sapply(strsplit(isoform_annotations_ccRCC3_tama$cell_barcodes, ","), length)
isoform_annotations_ccRCC4_tama$number_of_Cellbarcodes <- sapply(strsplit(isoform_annotations_ccRCC4_tama$cell_barcodes, ","), length)
isoform_annotations_ccRCC5_tama$number_of_Cellbarcodes <- sapply(strsplit(isoform_annotations_ccRCC5_tama$cell_barcodes, ","), length)
```


```{r NumberOfMatchingIsoforms2}
#UpsetPlot
isoform_annotations_normal_tama_3cells <- isoform_annotations_normal_tama[isoform_annotations_normal_tama$number_of_Cellbarcodes >= 3, ]
isoform_annotations_ccRCC2_tama_3cells <- isoform_annotations_ccRCC2_tama[isoform_annotations_ccRCC2_tama$number_of_Cellbarcodes >= 3, ]
isoform_annotations_ccRCC3_tama_3cells <- isoform_annotations_ccRCC3_tama[isoform_annotations_ccRCC3_tama$number_of_Cellbarcodes >= 3, ]
isoform_annotations_ccRCC4_tama_3cells <- isoform_annotations_ccRCC4_tama[isoform_annotations_ccRCC4_tama$number_of_Cellbarcodes >= 3, ]
isoform_annotations_ccRCC5_tama_3cells <- isoform_annotations_ccRCC5_tama[isoform_annotations_ccRCC5_tama$number_of_Cellbarcodes >= 3, ]

library(UpSetR)
IsoformIds_list_samples2 <- list(Normal = unique(isoform_annotations_normal_tama_3cells$TamaID), ccRCC2 = unique(isoform_annotations_ccRCC2_tama_3cells$TamaID), ccRCC3 = unique(isoform_annotations_ccRCC3_tama_3cells$TamaID), ccRCC4 = unique(isoform_annotations_ccRCC4_tama_3cells$TamaID), ccRCC5 = unique(isoform_annotations_ccRCC5_tama_3cells$TamaID))

IsoformMatching2 <- upset(fromList(IsoformIds_list_samples2), order.by = 'freq', nsets = 6, number.angles = 45, point.size = 3.5, line.size = 2, mainbar.y.label = "# Isoforms", sets.x.label = "Number of Isoforms", text.scale = 2)

IsoformMatching2

png("Figure4A.png", width = 16 * 300, height = 10 * 300, res = 300, bg = "transparent")
print(IsoformMatching2)
dev
```
sample2_4_5 <- intersect(intersect(isoform_annotations_ccRCC4_tama_3cells$TamaID, isoform_annotations_ccRCC5_tama_3cells$TamaID), isoform_annotations_ccRCC2_tama_3cells$TamaID)

sample2_4_5_uniq <- sample2_4_5[!sample2_4_5 %in% isoform_annotations_normal_tama_3cells$TamaID & !sample2_4_5 %in% isoform_annotations_ccRCC3_tama_3cells$TamaID] 

length(unique(isoform_annotations_ccRCC2_tama_3cells[isoform_annotations_ccRCC2_tama_3cells$TamaID %in% sample2_4_5_uniq, 'associated_gene']))
# novels
isoform_annotations_ccRCC2_tama_3cells[isoform_annotations_ccRCC2_tama_3cells$TamaID %in% sample2_4_5_uniq & isoform_annotations_ccRCC2_tama_3cells$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), ]


```{r readIsoforms}
ccRCC_Marker_Annotations <- read.table( '/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/CellTypes/ccRCC_CellType.tsv', sep='\t', header = TRUE)
ccRCC_Marker_Annotations$BC <- sapply(strsplit(ccRCC_Marker_Annotations$CellBarcode, "-"), function(x) x[1])

head(ccRCC_Marker_Annotations)
```

```{r readSeuratFiles}
seurat_annotations_ccRCC_2 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC2/scisoseq.seurat_info/scisoseq.annotated.info.csv', header = TRUE, sep='\t')

seurat_annotations_ccRCC_4 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC4/scisoseq.seurat_info/scisoseq.annotated.info.csv', header = TRUE, sep='\t')

seurat_annotations_ccRCC_5 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC5/scisoseq.seurat_info/scisoseq.annotated.info.csv', header = TRUE, sep='\t')

seurat_annotations_normal <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/Normal/scisoseq.seurat_info/scisoseq.annotated.info.csv', header = TRUE, sep='\t')

seurat_annotations_ccRCC3 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC3/scisoseq.seurat_info/scisoseq.annotated.info.csv', header = TRUE, sep='\t')

# assign Cells
seurat_annotations_ccRCC_2$Status <- ifelse(seurat_annotations_ccRCC_2$BC %in%  ccRCC_Marker_Annotations[ccRCC_Marker_Annotations$Sample == 'ccRCC2' & ccRCC_Marker_Annotations$CellType == 'ccRCC', 'BC'], 'ccRCC', 'non-ccRCC')

seurat_annotations_ccRCC_4$Status <- ifelse(seurat_annotations_ccRCC_4$BC %in%  ccRCC_Marker_Annotations[ccRCC_Marker_Annotations$Sample == 'ccRCC4' & ccRCC_Marker_Annotations$CellType == 'ccRCC', 'BC'], 'ccRCC', 'non-ccRCC')

seurat_annotations_ccRCC_4 <- seurat_annotations_ccRCC_4[seurat_annotations_ccRCC_4$BC %in% ccRCC_Marker_Annotations$BC, ]

seurat_annotations_ccRCC_5$Status <- ifelse(seurat_annotations_ccRCC_5$BC %in%  ccRCC_Marker_Annotations[ccRCC_Marker_Annotations$Sample == 'ccRCC5' & ccRCC_Marker_Annotations$CellType == 'ccRCC', 'BC'], 'ccRCC', 'non-ccRCC')

seurat_annotations_ccRCC3$Status <- 'non-ccRCC' 
seurat_annotations_normal$Status <- 'non-ccRCC' 
```

```{r mergeTama}
seurat_annotations_ccRCC_2_tama <- merge(isoform_annotations_ccRCC2_tama_3cells[,c(1,5,6,7,8,15,50,51,52)], seurat_annotations_ccRCC_2, by.y = 'pbid', by.x = 'isoform')

seurat_annotations_ccRCC_4_tama <- merge(isoform_annotations_ccRCC4_tama_3cells[,c(1,5,6,7,8,15,50,51,52)], seurat_annotations_ccRCC_4, by.y = 'pbid', by.x = 'isoform')

seurat_annotations_ccRCC_5_tama <- merge(isoform_annotations_ccRCC5_tama_3cells[,c(1,5,6,7,8,15,50,51,52)], seurat_annotations_ccRCC_5, by.y = 'pbid', by.x = 'isoform')

seurat_annotations_ccRCC_3_tama <- merge(isoform_annotations_ccRCC3_tama_3cells[,c(1,5,6,7,8,15,50,51,52)], seurat_annotations_ccRCC3, by.y = 'pbid', by.x = 'isoform')

seurat_annotations_normal_tama <- merge(isoform_annotations_normal_tama_3cells[,c(1,5,6,7,8,15,50,51,52)], seurat_annotations_normal, by.y = 'pbid', by.x = 'isoform')

```
length(unique(c(seurat_annotations_normal_tama$TamaID, seurat_annotations_ccRCC_2_tama$TamaID, seurat_annotations_ccRCC_4_tama$TamaID, seurat_annotations_ccRCC_5_tama$TamaID, seurat_annotations_ccRCC_3_tama$TamaID)))
[1] 311586

length(unique(c(seurat_annotations_normal_tama[seurat_annotations_normal_tama$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'], seurat_annotations_ccRCC_2_tama[seurat_annotations_ccRCC_2_tama$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'], seurat_annotations_ccRCC_4_tama[seurat_annotations_ccRCC_4_tama$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'], seurat_annotations_ccRCC_5_tama[seurat_annotations_ccRCC_5_tama$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'], seurat_annotations_ccRCC_3_tama[seurat_annotations_ccRCC_3_tama$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'])))
[1] 83009

%26 are novel
## Supplementary Fig 6
```{r SplitbyCA9PosAndNeg}
IsoformIds_CA9Pos_Neg <- list(NonccRCC = unique(c(seurat_annotations_normal_tama$TamaID, seurat_annotations_ccRCC_3_tama$TamaID, seurat_annotations_ccRCC_2_tama[seurat_annotations_ccRCC_2_tama$Status == 'non-ccRCC', 'TamaID'], seurat_annotations_ccRCC_4_tama[seurat_annotations_ccRCC_4_tama$Status == 'non-ccRCC', 'TamaID'], 
seurat_annotations_ccRCC_5_tama[seurat_annotations_ccRCC_5_tama$Status == 'non-ccRCC', 'TamaID'])),
ccRCC = unique(c(seurat_annotations_ccRCC_2_tama[seurat_annotations_ccRCC_2_tama$Status == 'ccRCC', 'TamaID'], seurat_annotations_ccRCC_4_tama[seurat_annotations_ccRCC_4_tama$Status == 'ccRCC', 'TamaID'], 
seurat_annotations_ccRCC_5_tama[seurat_annotations_ccRCC_5_tama$Status == 'ccRCC', 'TamaID'])))

ggVennDiagram(IsoformIds_CA9Pos_Neg, label_alpha = 0) + scale_fill_gradient(low="grey90",high = "red")


IsoformIds_CA9Pos_Neg2 <- list(
  Normal = unique(seurat_annotations_normal_tama$TamaID), 
ccRCC3 = unique(seurat_annotations_ccRCC_3_tama$TamaID), 
ccRCC2_NonccRCC = unique(seurat_annotations_ccRCC_2_tama[seurat_annotations_ccRCC_2_tama$Status == 'non-ccRCC', 'TamaID']), 
ccRCC4_NonccRCC = unique(seurat_annotations_ccRCC_4_tama[seurat_annotations_ccRCC_4_tama$Status == 'non-ccRCC', 'TamaID']), 
ccRCC5_NonccRCC = unique(seurat_annotations_ccRCC_5_tama[seurat_annotations_ccRCC_5_tama$Status == 'non-ccRCC', 'TamaID']),

ccRCC2_ccRCC = unique(seurat_annotations_ccRCC_2_tama[seurat_annotations_ccRCC_2_tama$Status == 'ccRCC', 'TamaID']), ccRCC4_ccRCC = unique(seurat_annotations_ccRCC_4_tama[seurat_annotations_ccRCC_4_tama$Status == 'ccRCC', 'TamaID']),
ccRCC5_ccRCC = unique(seurat_annotations_ccRCC_5_tama[seurat_annotations_ccRCC_5_tama$Status == 'ccRCC', 'TamaID']))

IsoformMatchingCA9 <- upset(fromList(IsoformIds_CA9Pos_Neg2), order.by = 'freq', nsets = 10, nintersects = 60, number.angles = 45, point.size = 3.5, line.size = 2, mainbar.y.label = "# Isoforms", sets.x.label = "Number of Isoforms", text.scale = 2)

IsoformMatchingCA9

png("Supplementary6.png", width = 20 * 400, height = 14 * 300, res = 300, bg = "transparent")
print(IsoformMatchingCA9)
dev.off()
```


# Supplementary Fig 7
```{r JaccardSimilarity}
jaccard_similarity_Novel_filtered <- function(isoAnnotation1, isoAnnotation2) {
   
    isoAnnotation1 <- isoAnnotation1[isoAnnotation1$number_of_Cellbarcodes >= 3, ]
    isoAnnotation2 <- isoAnnotation2[isoAnnotation2$number_of_Cellbarcodes >= 3, ]
  
    intersect1_2 <-length(intersect(isoAnnotation1[isoAnnotation1$structural_category %in% c('novel_not_in_catalog', 'novel_in_catalog'), 'TamaID'],isoAnnotation2[isoAnnotation2$structural_category %in% c('novel_not_in_catalog', 'novel_in_catalog'), 'TamaID']))
                        
  distinct1 <- length(isoAnnotation1[isoAnnotation1$structural_category %in% c('novel_not_in_catalog', 'novel_in_catalog'), 'TamaID']) - intersect1_2
  distinct2 <- length(isoAnnotation2[isoAnnotation2$structural_category %in% c('novel_not_in_catalog', 'novel_in_catalog'), 'TamaID']) - intersect1_2
  union1_2 <- intersect1_2 + distinct1 + distinct2
  similarity <- intersect1_2/union1_2
  return(similarity)
}

JC1_1 <- 1
JC1_2 <- jaccard_similarity_Novel_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC2_tama)
JC1_3 <- jaccard_similarity_Novel_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC3_tama)
JC1_4 <- jaccard_similarity_Novel_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC4_tama)
JC1_5 <- jaccard_similarity_Novel_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC5_tama)

JC2_2 <- 1
JC2_3 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC3_tama)
JC2_4 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC4_tama)
JC2_5 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC5_tama)

JC3_3 <- 1
JC3_4 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC4_tama)
JC3_5 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC5_tama)
JC4_4 <- 1
JC4_5 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC4_tama, isoform_annotations_ccRCC5_tama)
JC5_5 <- 1

jaccard_matrix <- matrix(0, nrow = 5, ncol = 5)
diag(jaccard_matrix) <- 1
jaccard_matrix[lower.tri(jaccard_matrix)] <- c(JC1_2, JC1_3, JC1_4, JC1_5, JC2_3, JC2_4, JC2_5, JC3_4, JC3_5, JC4_5)
jaccard_matrix[upper.tri(jaccard_matrix)] <- t(jaccard_matrix)[upper.tri(jaccard_matrix)]

# Generate the heatmap
library(pheatmap)
colnames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
rownames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
breaks <- seq(0, 0.4, length.out = 100)
colors <- colorRampPalette(c("blue", "red"))(99)

FigureSupp5A <- pheatmap(jaccard_matrix, 
         color = colors, 
         breaks = breaks,
         clustering_distance_rows = "euclidean", 
         clustering_distance_cols = "euclidean",
         display_numbers = TRUE, 
         number_color = "white", 
         cluster_rows = FALSE,
         cluster_cols = FALSE)

ggsave("Supplementary_Figure_7A.png", plot = FigureSupp5A, bg = "transparent", width = 5, height = 4)
```

```{r 3Allcellssimilarity}
jaccard_similarity_filtered <- function(isoAnnotation1, isoAnnotation2) {
   
    isoAnnotation1 <- isoAnnotation1[isoAnnotation1$number_of_Cellbarcodes >= 3, ]
    isoAnnotation2 <- isoAnnotation2[isoAnnotation2$number_of_Cellbarcodes >= 3, ]
  
    intersect1_2 <-length(intersect(isoAnnotation1$TamaID,isoAnnotation2$TamaID))
                        
  distinct1 <- length(isoAnnotation1$TamaID) - intersect1_2
  distinct2 <- length(isoAnnotation2$TamaID) - intersect1_2
  union1_2 <- intersect1_2 + distinct1 + distinct2
  similarity <- intersect1_2/union1_2
  return(similarity)
}


JC1_1 <- 1
JC1_2 <- jaccard_similarity_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC2_tama)
JC1_3 <- jaccard_similarity_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC3_tama)
JC1_4 <- jaccard_similarity_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC4_tama)
JC1_5 <- jaccard_similarity_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC5_tama)

JC2_2 <- 1
JC2_3 <- jaccard_similarity_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC3_tama)
JC2_4 <- jaccard_similarity_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC4_tama)
JC2_5 <- jaccard_similarity_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC5_tama)

JC3_3 <- 1
JC3_4 <- jaccard_similarity_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC4_tama)
JC3_5 <- jaccard_similarity_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC5_tama)
JC4_4 <- 1
JC4_5 <- jaccard_similarity_filtered(isoform_annotations_ccRCC4_tama, isoform_annotations_ccRCC5_tama)
JC5_5 <- 1

jaccard_matrix <- matrix(0, nrow = 5, ncol = 5)
diag(jaccard_matrix) <- 1
jaccard_matrix[lower.tri(jaccard_matrix)] <- c(JC1_2, JC1_3, JC1_4, JC1_5, JC2_3, JC2_4, JC2_5, JC3_4, JC3_5, JC4_5)
jaccard_matrix[upper.tri(jaccard_matrix)] <- t(jaccard_matrix)[upper.tri(jaccard_matrix)]

# Generate the heatmap
library(pheatmap)
colnames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
rownames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
breaks <- seq(0, 0.4, length.out = 100)
colors <- colorRampPalette(c("blue", "red"))(99)

figure5A_ALL <- pheatmap(jaccard_matrix, 
         color = colors, 
         breaks = breaks,
         clustering_distance_rows = "euclidean", 
         clustering_distance_cols = "euclidean",
         display_numbers = TRUE, 
         number_color = "white", 
         cluster_rows = FALSE,
         cluster_cols = FALSE)


ggsave("Supplementary_Figure_7A_all.png", plot = figure5A_ALL, bg = "transparent", width = 5, height = 4)

```



```{r 3cellssimilarityFSM}
jaccard_similarity_FSM_filtered <-  function(isoAnnotation1, isoAnnotation2) {
   
    isoAnnotation1 <- isoAnnotation1[isoAnnotation1$number_of_Cellbarcodes >= 3, ]
    isoAnnotation2 <- isoAnnotation2[isoAnnotation2$number_of_Cellbarcodes >= 3, ]
  
    intersect1_2 <-length(intersect(isoAnnotation1[isoAnnotation1$structural_category %in% c('full-splice_match'), 'TamaID'],isoAnnotation2[isoAnnotation2$structural_category %in% c('full-splice_match'), 'TamaID']))
                        
  distinct1 <- length(isoAnnotation1[isoAnnotation1$structural_category %in% c('full-splice_match'), 'TamaID']) - intersect1_2
  distinct2 <- length(isoAnnotation2[isoAnnotation2$structural_category %in% c('full-splice_match'), 'TamaID']) - intersect1_2
  union1_2 <- intersect1_2 + distinct1 + distinct2
  similarity <- intersect1_2/union1_2
  return(similarity)
}


JC1_1 <- 1
JC1_2 <- jaccard_similarity_FSM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC2_tama)
JC1_3 <- jaccard_similarity_FSM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC3_tama)
JC1_4 <- jaccard_similarity_FSM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC4_tama)
JC1_5 <- jaccard_similarity_FSM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC5_tama)

JC2_2 <- 1
JC2_3 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC3_tama)
JC2_4 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC4_tama)
JC2_5 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC5_tama)

JC3_3 <- 1
JC3_4 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC4_tama)
JC3_5 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC5_tama)
JC4_4 <- 1
JC4_5 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC4_tama, isoform_annotations_ccRCC5_tama)
JC5_5 <- 1

jaccard_matrix <- matrix(0, nrow = 5, ncol = 5)
diag(jaccard_matrix) <- 1
jaccard_matrix[lower.tri(jaccard_matrix)] <- c(JC1_2, JC1_3, JC1_4, JC1_5, JC2_3, JC2_4, JC2_5, JC3_4, JC3_5, JC4_5)
jaccard_matrix[upper.tri(jaccard_matrix)] <- t(jaccard_matrix)[upper.tri(jaccard_matrix)]

# Generate the heatmap
library(pheatmap)
colnames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
rownames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
breaks <- seq(0.2, 0.4, length.out = 100)
colors <- colorRampPalette(c("blue", "red"))(99)

figure5a_fsm <- pheatmap(jaccard_matrix, 
         color = colors, 
         breaks = breaks,
         clustering_distance_rows = "euclidean", 
         clustering_distance_cols = "euclidean",
         display_numbers = TRUE, 
         number_color = "white", 
         cluster_rows = FALSE,
         cluster_cols = FALSE)


ggsave("Supplementary_Figure_7A_fsm.png", plot = figure5a_fsm, bg = "transparent", width = 5, height = 4)
```


```{r 3cellssimilarityISM}
jaccard_similarity_ISM_filtered <-  function(isoAnnotation1, isoAnnotation2) {
   
    isoAnnotation1 <- isoAnnotation1[isoAnnotation1$number_of_Cellbarcodes >= 3, ]
    isoAnnotation2 <- isoAnnotation2[isoAnnotation2$number_of_Cellbarcodes >= 3, ]
  
    intersect1_2 <-length(intersect(isoAnnotation1[isoAnnotation1$structural_category %in% c('incomplete-splice_match'), 'TamaID'],isoAnnotation2[isoAnnotation2$structural_category %in% c('incomplete-splice_match'), 'TamaID']))
                        
  distinct1 <- length(isoAnnotation1[isoAnnotation1$structural_category %in% c('incomplete-splice_match'), 'TamaID']) - intersect1_2
  distinct2 <- length(isoAnnotation2[isoAnnotation2$structural_category %in% c('incomplete-splice_match'), 'TamaID']) - intersect1_2
  union1_2 <- intersect1_2 + distinct1 + distinct2
  similarity <- intersect1_2/union1_2
  return(similarity)
}

  
JC1_1 <- 1
JC1_2 <- jaccard_similarity_ISM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC2_tama)
JC1_3 <- jaccard_similarity_ISM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC3_tama)
JC1_4 <- jaccard_similarity_ISM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC4_tama)
JC1_5 <- jaccard_similarity_ISM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC5_tama)

JC2_2 <- 1
JC2_3 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC3_tama)
JC2_4 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC4_tama)
JC2_5 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC5_tama)

JC3_3 <- 1
JC3_4 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC4_tama)
JC3_5 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC5_tama)
JC4_4 <- 1
JC4_5 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC4_tama, isoform_annotations_ccRCC5_tama)
JC5_5 <- 1

jaccard_matrix <- matrix(0, nrow = 5, ncol = 5)
diag(jaccard_matrix) <- 1
jaccard_matrix[lower.tri(jaccard_matrix)] <- c(JC1_2, JC1_3, JC1_4, JC1_5, JC2_3, JC2_4, JC2_5, JC3_4, JC3_5, JC4_5)
jaccard_matrix[upper.tri(jaccard_matrix)] <- t(jaccard_matrix)[upper.tri(jaccard_matrix)]

# Generate the heatmap
library(pheatmap)
colnames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
rownames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
breaks <- seq(0.1, 0.3, length.out = 100)
colors <- colorRampPalette(c("blue", "red"))(99)

figure5a_ism <- pheatmap(jaccard_matrix, 
         color = colors, 
         breaks = breaks,
         clustering_distance_rows = "euclidean", 
         clustering_distance_cols = "euclidean",
         display_numbers = TRUE, 
         number_color = "white", 
         cluster_rows = FALSE,
         cluster_cols = FALSE)



ggsave("Supplementary_Figure_7A_ism.png", plot = figure5a_ism, bg = "transparent", width = 5, height = 4)
```

# Figure 4B

```{r isoformmatch}
AllMatchingIDs <- intersect(intersect(intersect(intersect(isoform_annotations_ccRCC2_tama_3cells$TamaID, isoform_annotations_ccRCC3_tama_3cells$TamaID), isoform_annotations_ccRCC4_tama_3cells$TamaID), isoform_annotations_ccRCC5_tama_3cells$TamaID), isoform_annotations_normal_tama_3cells$TamaID)
```

```{r readIsoMatches}
CellNumberInNormal <- as.data.frame(isoform_annotations_normal_tama_3cells[isoform_annotations_normal_tama_3cells$TamaID %in% AllMatchingIDs, 'number_of_Cellbarcodes'])
CellNumberInccRCC2 <- as.data.frame(isoform_annotations_ccRCC2_tama_3cells[isoform_annotations_ccRCC2_tama_3cells$TamaID %in% AllMatchingIDs, 'number_of_Cellbarcodes'])
CellNumberInccRCC3 <- as.data.frame(isoform_annotations_ccRCC3_tama_3cells[isoform_annotations_ccRCC3_tama_3cells$TamaID %in% AllMatchingIDs, 'number_of_Cellbarcodes'])
CellNumberInccRCC4 <- as.data.frame(isoform_annotations_ccRCC4_tama_3cells[isoform_annotations_ccRCC4_tama_3cells$TamaID %in% AllMatchingIDs, 'number_of_Cellbarcodes'])
CellNumberInccRCC5 <- as.data.frame(isoform_annotations_ccRCC5_tama_3cells[isoform_annotations_ccRCC5_tama_3cells$TamaID %in% AllMatchingIDs, 'number_of_Cellbarcodes'])

CellNumberInNormal_nonmatch <- as.data.frame(isoform_annotations_normal_tama_3cells[!isoform_annotations_normal_tama_3cells$TamaID %in% AllMatchingIDs, 'number_of_Cellbarcodes'])
CellNumberInccRCC2_nonmatch <- as.data.frame(isoform_annotations_ccRCC2_tama_3cells[!isoform_annotations_ccRCC2_tama_3cells$TamaID %in% AllMatchingIDs, 'number_of_Cellbarcodes'])
CellNumberInccRCC3_nonmatch <- as.data.frame(isoform_annotations_ccRCC3_tama_3cells[!isoform_annotations_ccRCC3_tama_3cells$TamaID %in% AllMatchingIDs, 'number_of_Cellbarcodes'])
CellNumberInccRCC4_nonmatch <- as.data.frame(isoform_annotations_ccRCC4_tama_3cells[!isoform_annotations_ccRCC4_tama_3cells$TamaID %in% AllMatchingIDs, 'number_of_Cellbarcodes'])
CellNumberInccRCC5_nonmatch <- as.data.frame(isoform_annotations_ccRCC5_tama_3cells[!isoform_annotations_ccRCC5_tama_3cells$TamaID %in% AllMatchingIDs, 'number_of_Cellbarcodes'])



CellNumberInNormal$Group <- 'Normal'
CellNumberInccRCC2$Group <- 'ccRCC2'
CellNumberInccRCC3$Group <- 'ccRCC3'
CellNumberInccRCC4$Group <- 'ccRCC4'
CellNumberInccRCC5$Group <- 'ccRCC5'

CellNumberInNormal$Status <- 'Matched'
CellNumberInccRCC2$Status <- 'Matched'
CellNumberInccRCC3$Status <- 'Matched'
CellNumberInccRCC4$Status <- 'Matched'
CellNumberInccRCC5$Status <- 'Matched'


CellNumberInNormal_nonmatch$Group <- 'Normal'
CellNumberInccRCC2_nonmatch$Group <- 'ccRCC2'
CellNumberInccRCC3_nonmatch$Group <- 'ccRCC3'
CellNumberInccRCC4_nonmatch$Group <- 'ccRCC4'
CellNumberInccRCC5_nonmatch$Group <- 'ccRCC5'

CellNumberInNormal_nonmatch$Status <- 'UnMatched'
CellNumberInccRCC2_nonmatch$Status <- 'UnMatched'
CellNumberInccRCC3_nonmatch$Status <- 'UnMatched'
CellNumberInccRCC4_nonmatch$Status <- 'UnMatched'
CellNumberInccRCC5_nonmatch$Status <- 'UnMatched'

colnames(CellNumberInccRCC5) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC4) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC3) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC2) <- c('CellNumber', 'Group', 'Status')


colnames(CellNumberInNormal) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC4_nonmatch) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC3_nonmatch) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC2_nonmatch) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInNormal_nonmatch) <- c('CellNumber', 'Group', 'Status')

colnames(CellNumberInccRCC5_nonmatch) <- c('CellNumber', 'Group', 'Status')
All_Matched_cellNumber <- rbind(CellNumberInNormal, CellNumberInccRCC2, CellNumberInccRCC3, CellNumberInccRCC4, CellNumberInccRCC5)
colnames(All_Matched_cellNumber) <- c('CellNumber', 'Group', 'Status')
All_UnMatched_cellNumber <-  rbind(CellNumberInNormal_nonmatch, CellNumberInccRCC2_nonmatch, CellNumberInccRCC4_nonmatch,CellNumberInccRCC3_nonmatch, CellNumberInccRCC5_nonmatch)
colnames(All_UnMatched_cellNumber) <- c('CellNumber', 'Group', 'Status')
All_Matched_UnMatched <- rbind(All_Matched_cellNumber, All_UnMatched_cellNumber)
head(All_Matched_UnMatched)
```


```{r DensityPlot}
figure5bsupp <- ggplot(All_Matched_UnMatched, aes(log2(CellNumber), fill=Status)) + geom_density(alpha=0.5) + facet_wrap(~Group) + theme_bw()
figure5bsupp
mean(All_Matched_cellNumber$CellNumber)
#[1] 33.87988
mean(All_UnMatched_cellNumber$CellNumber)
#[1] 8.134875
ggsave("Figure_6B.png", plot = figure5bsupp, bg = "transparent", width = 15, height = 10)
```
