---
title: "Manuscript_1"
author: "Tülay Karakulak"
date: "2024-09-02"
output: html_document
---

# Codes to create Figure 1, Figure 2, and Supplementary Figure 1
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r required_libs, message=FALSE, warning=FALSE, include=FALSE}
library(Seurat)
library(patchwork)
library(dplyr)
library(ggplot2)
library(knitr)
library(rmarkdown)
library(harmony)
library(Matrix)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(ggpubr)
set.seed(12345)
```

## read Filtered Files
```{r readSeuratObjs}
# Isoform Level
healthy_isoform_subset <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/SeuratObj_filtered/Isoform_Level/healthy_isoform_subset.RDS')
ccRCC_2_isoform_subset <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/SeuratObj_filtered/Isoform_Level/ccRCC_2_isoform_subset.RDS')
ccRCC_3_isoform_subset <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/SeuratObj_filtered/Isoform_Level/ccRCC_3_isoform_subset.RDS')
ccRCC_4_isoform_subset <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/SeuratObj_filtered/Isoform_Level/ccRCC_4_isoform_subset.RDS')
ccRCC_5_isoform_subset <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/SeuratObj_filtered/Isoform_Level/ccRCC_5_isoform_subset.RDS')

## Gene Level
healthy_gene_subset <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/SeuratObj_filtered/Gene_Level/healthy_gene_subset.RDS')
ccRCC_2_gene_subset <- readRDS('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/SeuratObj_filtered/Gene_Level/ccRCC_2_gene_subset.RDS')
ccRCC_3_gene_subset <- readRDS('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/SeuratObj_filtered/Gene_Level/ccRCC_3_gene_subset.RDS')
ccRCC_4_gene_subset <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/SeuratObj_filtered/Gene_Level/ccRCC_4_gene_subset.RDS')
ccRCC_5_gene_subset <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/SeuratObj_filtered/Gene_Level/ccRCC_5_gene_subset.RDS')
```


```{r readSciSeqFiles}
scisoseq_file_normal_filtered2 <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/scisoseq_files_filtered/scisoseq_file_normal_filtered.RDS') # 101477 # 390 cells
scisoseq_file_ccRCC2_filtered2 <- readRDS( '//Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/scisoseq_files_filtered/scisoseq_file_ccRCC2_filtered.RDS') # 96847 # 334 cells
scisoseq_file_ccRCC3_filtered2 <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/scisoseq_files_filtered/scisoseq_file_ccRCC3_filtered.RDS') # 56384 # 272 cells
scisoseq_file_ccRCC4_filtered2 <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/scisoseq_files_filtered/scisoseq_file_ccRCC4_filtered.RDS') # 82847 # 1016 cells
scisoseq_file_ccRCC5_filtered2 <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/scisoseq_files_filtered/scisoseq_file_ccRCC5_filtered.RDS') # 91132 # 366 cells
```


```{r readIsoformAnnotationFiles}
isoform_annotations_normal_filtered_w_CellNumber <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/isoform_annotation_files_filtered/isoform_annotations_normal_filtered_w_CellNumber.RDS')
isoform_annotations_ccRCC_2_filtered_w_CellNumber <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/isoform_annotation_files_filtered/isoform_annotations_ccRCC_2_filtered_w_CellNumber.RDS')
isoform_annotations_ccRCC_3_filtered_w_CellNumber <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/isoform_annotation_files_filtered/isoform_annotations_ccRCC_3_filtered_w_CellNumber.RDS')
isoform_annotations_ccRCC_4_filtered_w_CellNumber <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/isoform_annotation_files_filtered/isoform_annotations_ccRCC_4_filtered_w_CellNumber.RDS')
isoform_annotations_ccRCC_5_filtered_w_CellNumber <- readRDS('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_Codes_Supplementaries/InputFiles/isoform_annotation_files_filtered/isoform_annotations_ccRCC_5_filtered_w_CellNumber.RDS')
```


# Data Analysis for the Manuscript Figures
## Figure 1B 
```{r SupplementaryFig1A}
kidney.combined_big <- merge(healthy_gene_subset, y = c(ccRCC_2_gene_subset, ccRCC_3_gene_subset, ccRCC_4_gene_subset, ccRCC_5_gene_subset), add.cell.ids = c("healthy","ccRCC_2", "ccRCC_3", "ccRCC_4", "ccRCC_5"), project = "ccRCC")
kidney.combined_big_isoform <- merge(healthy_isoform_subset, y = c(ccRCC_2_isoform_subset ,ccRCC_3_isoform_subset, ccRCC_4_isoform_subset, ccRCC_5_isoform_subset), add.cell.ids = c("healthy","ccRCC_2", "ccRCC_3", "ccRCC_4", "ccRCC_5"), project = "ccRCC")


a <- VlnPlot(kidney.combined_big, features = c("nFeature_RNA", "nCount_RNA"), ncol = 3)
b <- VlnPlot(kidney.combined_big_isoform, features = c("nFeature_RNA", "nCount_RNA"), ncol = 3)
a
b
#ggsave("/Figures/1B_Gene.tiff", plot = c, bg = "transparent", width = 8, height = 5, dpi=300)
#ggsave("/Figures/1B_Transcript.tiff", plot = b, bg = "transparent", width = 8, height = 5, dpi=300)
```


### Supplementary Figure 1A
```{r SupplementaryFig1B}
cellNumbers <- c(390, 334, 272, 1016, 366)

kidney.combined2 <- list(healthy = healthy_gene_subset, ccRCC_2 = ccRCC_2_gene_subset,  ccRCC_3 = ccRCC_3_gene_subset,  ccRCC_4 = ccRCC_4_gene_subset,  ccRCC_5 = ccRCC_5_gene_subset)

healthy_UMI <- mean(kidney.combined2$healthy@meta.data$nCount_RNA)
ccRCC2_UMI <-  mean(kidney.combined2$ccRCC_2@meta.data$nCount_RNA)
ccRCC3_UMI <-  mean(kidney.combined2$ccRCC_3@meta.data$nCount_RNA)
ccRCC4_UMI <-  mean(kidney.combined2$ccRCC_4@meta.data$nCount_RNA)
ccRCC5_UMI <-  mean(kidney.combined2$ccRCC_5@meta.data$nCount_RNA)

ReadMeanNumbers <- c(healthy_UMI, ccRCC2_UMI, ccRCC3_UMI, ccRCC4_UMI, ccRCC5_UMI)

CellReadNumbers <- data.frame(CellNumbers = cellNumbers, ReadMeanNumbers=ReadMeanNumbers)
rownames(CellReadNumbers) <- c('Healthy', 'ccRCC2', 'ccRCC3', 'ccRCC4', 'ccRCC5')


CellReadMeanNumber <- ggscatter(CellReadNumbers, x = "CellNumbers", y = "ReadMeanNumbers",
          add = "reg.line",                                 # Add regression line
          conf.int = TRUE,                                  # Add confidence interval
          add.params = list(color = "blue",
                            fill = "lightgray"))+
stat_cor(method = "pearson", label.x = 6, label.y = 15)       #   # Add correlation coefficient

CellReadMeanNumber
#ggsave("Supplementary_1A_CellReadMeanNumber.tiff", plot = CellReadMeanNumber, bg = "transparent", width = 8, height = 5, dpi=600)
```


```{r QC}
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
kidney.combined_big2[["percent.mt"]] <- PercentageFeatureSet(kidney.combined_big2, pattern = "^MT-")
kidney.combined_big2[["percent.ribo"]] <- PercentageFeatureSet(kidney.combined_big2, pattern = "^RP[SL]")

head(kidney.combined_big2@meta.data, 5) # number of unique genes and total molecules 

VlnPlot(kidney.combined_big2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r featurescatter}
plot1 <- FeatureScatter(kidney.combined_big2, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1
plot2 <- FeatureScatter(kidney.combined_big_isoform, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2
```

# Number of Transcript and Number of Novel Isoform details in the Manuscript
```{r CheckMeans}
# Check how many isoforms are detected on average
mean(dim(isoform_annotations_ccRCC_2_filtered_w_CellNumber)[1], dim(isoform_annotations_ccRCC_3_filtered_w_CellNumber)[1], dim(isoform_annotations_ccRCC_4_filtered_w_CellNumber)[1], dim(isoform_annotations_ccRCC_5_filtered_w_CellNumber)[1], dim(isoform_annotations_normal_filtered_w_CellNumber)[1]) # 97294

mean(dim(isoform_annotations_ccRCC_2_filtered_w_CellNumber[isoform_annotations_ccRCC_2_filtered_w_CellNumber$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),])[1], dim(isoform_annotations_ccRCC_3_filtered_w_CellNumber[isoform_annotations_ccRCC_3_filtered_w_CellNumber$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),])[1], dim(isoform_annotations_ccRCC_4_filtered_w_CellNumber[isoform_annotations_ccRCC_4_filtered_w_CellNumber$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),])[1], dim(isoform_annotations_ccRCC_5_filtered_w_CellNumber[isoform_annotations_ccRCC_5_filtered_w_CellNumber$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),])[1], dim(isoform_annotations_normal_filtered_w_CellNumber[isoform_annotations_normal_filtered_w_CellNumber$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),])[1]) # 38188
```

Percentage of Novel Isoforms: 38188/97294*100

```{r mergeAllSamples}
isoform_annotations_normal_filtered_w_CellNumber$Group <- 'Normal'
isoform_annotations_ccRCC_2_filtered_w_CellNumber$Group <- 'ccRCC2'
isoform_annotations_ccRCC_3_filtered_w_CellNumber$Group <- 'ccRCC3'
isoform_annotations_ccRCC_4_filtered_w_CellNumber$Group <- 'ccRCC4'
isoform_annotations_ccRCC_5_filtered_w_CellNumber$Group <- 'ccRCC5'
isoform_annotations <- rbind(isoform_annotations_normal_filtered_w_CellNumber, isoform_annotations_ccRCC_2_filtered_w_CellNumber, isoform_annotations_ccRCC_3_filtered_w_CellNumber, isoform_annotations_ccRCC_4_filtered_w_CellNumber, isoform_annotations_ccRCC_5_filtered_w_CellNumber)
```

### Figure 1E
```{r Figure1H}
# proportion of main categories
isoform_annotations$structural_category[isoform_annotations$structural_category %in% c('antisense', 'intergenic', 'fusion', 'moreJunctions', 'genic')] <- 'Other'

Structural_Category <- isoform_annotations %>%
    group_by(Group, structural_category) %>%
    summarise(count = n()) %>%
    mutate(total = sum(count)) %>%
    mutate(proportion = count/total) %>% ungroup()

color_palette <- viridis(5, option = "D")

c <- ggplot(Structural_Category, aes(x = Group, y = proportion, fill = structural_category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Proportion", x = "Sample", title = "Proportion of Transcript Categories") +
  scale_fill_manual(values = color_palette) + 
  theme_classic() +
  theme(legend.position="right", plot.background = element_rect(fill = NA, color = NA))
c

# Save the plot with a transparent background
#ggsave("Figures/Figure1E.png", plot = c, bg = "transparent", width = 8, height = 5, dpi=300)
```


```{r checkMeanProportions}
# mean novel isoforms
mean(as.vector(Structural_Category[Structural_Category$structural_category == 'novel_not_in_catalog', 'proportion'] + Structural_Category[Structural_Category$structural_category == 'novel_in_catalog', 'proportion'])$proportion) # number of Novel Isoforms # 0.3566599

# mean incomplete-splice-match
mean(Structural_Category[Structural_Category$structural_category == 'incomplete-splice_match', 'proportion']$proportion)
# 0.09387169
```

Average Novel isoform across Samples: 0.3566599

### Figure 1H
```{r Figure_1H}
color_palette2 <- viridis(5, option = "D")
TranscriptLength_ALL <- ggplot(isoform_annotations, aes(x = structural_category, y = length, fill=structural_category)) +
  geom_boxplot() +
  labs(y = "Length", x = "Groups", title = "Transcript Lengths") +
   scale_fill_manual(values = color_palette2) + 
    theme_classic() +
    facet_wrap(~Group, scales="free_x") +
  #facet_wrap(~Group, scales="free_x") +
  theme(legend.position="right") + 
  theme(axis.text.x = element_blank())

TranscriptLength_ALL
#ggsave("Figure_1H.png", plot = TranscriptLength_ALL, bg = "transparent", width = 8, height = 5, dpi=300)
```

Statistical Test
t.test(isoform_annotations[isoform_annotations$structural_category == 'full-splice_match', 'length'], isoform_annotations[isoform_annotations$structural_category == 'novel_in_catalog', 'length'])

t.test(isoform_annotations[isoform_annotations$structural_category %in% c('full-splice_match', 'novel_in_catalog'), 'length'], isoform_annotations[isoform_annotations$structural_category == 'incomplete-splice_match', 'length'])

	Welch Two Sample t-test

data:  isoform_annotations[isoform_annotations$structural_category %in% c("full-splice_match", "novel_in_catalog"), "length"] and isoform_annotations[isoform_annotations$structural_category == "incomplete-splice_match", "length"]
t = 107.34, df = 55370, p-value < 2.2e-16
alternative hypothesis: true difference in means is not equal to 0
95 percent confidence interval:
 343.1608 355.9258
sample estimates:
mean of x mean of y 
1337.6398  988.0966 

## Supplementary Fig 1B
```{r Supplementary_Figure2}
# proportion of main categories
Structural_sub_Category <- isoform_annotations %>%
    group_by(Group, structural_category, subcategory) %>%
    summarise(count = n()) %>%
    mutate(total = sum(count)) %>%
    mutate(proportion = count/total) %>% ungroup()

color_palette <- viridis(14, option = "D")

c <- ggplot(Structural_sub_Category, aes(x = structural_category, y = proportion, fill = subcategory)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Proportion", x = "Sample", title = "Proportion of Transcript SubCategories") +
  scale_fill_manual(values = color_palette) + 
  theme_bw() +
  facet_wrap(~Group, scales="free_x", nrow=1) +
  theme(legend.position="right", axis.text.x = element_text(angle = 90, hjust = 1))
c

#ggsave("Supplementary_Figure_1B.png", plot = c, bg = "transparent", width = 12, height = 5)
```
isoform_annotations %>%
    group_by(structural_category, subcategory) %>%
    summarise(count = n()) %>%
    mutate(total = sum(count)) %>%
    mutate(proportion = count/total) %>% ungroup()

## Supplementary Table 1 -  isoforms found more than 30% of cells of a sample
```{r SupplementaryTable1}
# Normal has 390 cells
Suppl_Table1_Normal <- isoform_annotations[isoform_annotations$CellNumber >= 117 & isoform_annotations$Group == 'Normal' & isoform_annotations$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),] %>% arrange(desc(CellNumber)) %>% dplyr::select(isoform, chrom, strand, length, exons, structural_category, structural_category, associated_gene.x, ref_length, ref_exons, diff_to_gene_TSS, diff_to_gene_TTS, subcategory, RTS_stage, FL, perc_A_downstream_TTS, dist_to_cage_peak, within_cage_peak, polyA_dist, fl_assoc,Group, CellNumber)

# ccRCC2 has 334 cells
Suppl_Table1_ccRCC2 <- isoform_annotations[isoform_annotations$CellNumber >= 100.2 & isoform_annotations$Group == 'ccRCC2' & isoform_annotations$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),] %>% arrange(desc(CellNumber)) %>% dplyr::select(isoform, chrom, strand, length, exons, structural_category, structural_category, associated_gene.x, ref_length, ref_exons, diff_to_gene_TSS, diff_to_gene_TTS, subcategory, RTS_stage, FL, perc_A_downstream_TTS, dist_to_cage_peak, within_cage_peak, polyA_dist, fl_assoc,Group, CellNumber)

# ccRCC3 has 272 cells
Suppl_Table1_ccRCC3 <- isoform_annotations[isoform_annotations$CellNumber >= 81.6 & isoform_annotations$Group == 'ccRCC3' & isoform_annotations$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),] %>% arrange(desc(CellNumber)) %>% dplyr::select(isoform, chrom, strand, length, exons, structural_category, structural_category, associated_gene.x, ref_length, ref_exons, diff_to_gene_TSS, diff_to_gene_TTS, subcategory, RTS_stage, FL, perc_A_downstream_TTS, dist_to_cage_peak, within_cage_peak, polyA_dist, fl_assoc,Group, CellNumber)

# ccRCC4 has 1016 cells
Suppl_Table1_ccRCC4 <- isoform_annotations[isoform_annotations$CellNumber >= 304.8 & isoform_annotations$Group == 'ccRCC4' & isoform_annotations$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),] %>% arrange(desc(CellNumber)) %>% dplyr::select(isoform, chrom, strand, length, exons, structural_category, structural_category, associated_gene.x, ref_length, ref_exons, diff_to_gene_TSS, diff_to_gene_TTS, subcategory, RTS_stage, FL, perc_A_downstream_TTS, dist_to_cage_peak, within_cage_peak, polyA_dist, fl_assoc,Group, CellNumber)

# ccRCC4 has 366 cells
Suppl_Table1_ccRCC5 <- isoform_annotations[isoform_annotations$CellNumber >= 109.8 & isoform_annotations$Group == 'ccRCC5' & isoform_annotations$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),] %>% arrange(desc(CellNumber)) %>% dplyr::select(isoform, chrom, strand, length, exons, structural_category, structural_category, associated_gene.x, ref_length, ref_exons, diff_to_gene_TSS, diff_to_gene_TTS, subcategory, RTS_stage, FL, perc_A_downstream_TTS, dist_to_cage_peak, within_cage_peak, polyA_dist, fl_assoc, Group, CellNumber)

SupplementaryTable1 <- rbind(Suppl_Table1_ccRCC5, Suppl_Table1_ccRCC4, Suppl_Table1_ccRCC3, Suppl_Table1_ccRCC2, Suppl_Table1_Normal)
dim(SupplementaryTable1)
# save the file
#write.table(SupplementaryTable1, 'SupplementaryTable1.tsv', row.names = FALSE, quote = FALSE)
```
total of 686 novel isoforms found in more than 30% of the cells of a sample.


### check the percentage of number of novel isoforms found in 1 cells
dim(isoform_annotations[isoform_annotations$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),])
[1] 157654     52

dim(isoform_annotations[isoform_annotations$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog') & isoform_annotations$CellNumber == 1,])
[1] 99758    52
99758/157654*100
63.27654

intersect of novel genes
intersect(SupplementaryTable1[SupplementaryTable1$Group == 'ccRCC2', 'associated_gene.x'], SupplementaryTable1[SupplementaryTable1$Group == 'ccRCC5', 'associated_gene.x'])


## Figure 1F and Figure 1G
```{r Fig1E_SupplementaryFig1D}
# Figure 1E 
# after filtering
isoform_annotations_IsoGene <- isoform_annotations %>% group_by(Group, associated_gene.x) %>% summarise(NIsoform = n())
isoform_annotations_IsoGene_novel <- isoform_annotations %>% dplyr::filter(structural_category %in% c('novel_in_catalog','novel_not_in_catalog')) %>% group_by(Group, associated_gene.x) %>% summarise(NIsoform = n())

isoform_annotations_IsoGene$bins <- cut(isoform_annotations_IsoGene$NIsoform, 
                                         breaks=c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, Inf), labels=c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", ">10"),  right=FALSE)


GeneIsoformRatio <- isoform_annotations_IsoGene %>% group_by(Group,bins) %>% summarise(NumberOfGenes = n())

isoform_annotations_IsoGene_novel$bins <- cut(isoform_annotations_IsoGene_novel$NIsoform, 
                                         breaks=c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, Inf), labels=c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", ">10"),  right=FALSE)

GeneIsoformRatio_novel <- isoform_annotations_IsoGene_novel %>% group_by(Group,bins) %>% summarise(NumberOfGenes = n())

# Figure 1E
p2 <- ggplot(GeneIsoformRatio, aes(x = bins, y = NumberOfGenes)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Number of Gene", x = "Number of Transcript", title = "The number of Isoforms Per Gene") +
  theme_classic() +
  facet_wrap(~Group, scales="free_x") +
  #theme(legend.position="bottom")
  theme(legend.position="right") + 
  theme(axis.text.x = element_blank())

p2

p3 <- ggplot(GeneIsoformRatio_novel, aes(x = bins, y = NumberOfGenes)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Number of Gene", x = "Number of Transcript", title = "The number of Novel Isoforms Per Gene") +
  theme_classic() +
  facet_wrap(~Group, scales="free_x") +
  #theme(legend.position="bottom")
  theme(legend.position="right") + 
  theme(axis.text.x = element_blank())

p3

#ggsave("Figure1F_All.png", plot = p2, bg = "transparent", width = 6, height = 5)
#ggsave("Figure1G_Novel.png", plot = p3, bg = "transparent", width = 6, height = 5)
```


```{r percentagesOfGeneIsoformRatios}
GeneIsoformRatio[GeneIsoformRatio$Group == 'Normal', 'NumberOfGenes']/sum(GeneIsoformRatio[GeneIsoformRatio$Group == 'Normal', 'NumberOfGenes']$NumberOfGenes)*100

GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC2', 'NumberOfGenes']/sum(GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC2', 'NumberOfGenes']$NumberOfGenes)*100

GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC3', 'NumberOfGenes']/sum(GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC3', 'NumberOfGenes']$NumberOfGenes)*100

GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC4', 'NumberOfGenes']/sum(GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC4', 'NumberOfGenes']$NumberOfGenes)*100

GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC5', 'NumberOfGenes']/sum(GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC5', 'NumberOfGenes']$NumberOfGenes)*100


GeneIsoformRatio_novel[GeneIsoformRatio_novel$Group == 'Normal', 'NumberOfGenes']/sum(GeneIsoformRatio[GeneIsoformRatio$Group == 'Normal', 'NumberOfGenes']$NumberOfGenes)*100

GeneIsoformRatio_novel[GeneIsoformRatio_novel$Group == 'ccRCC2', 'NumberOfGenes']/sum(GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC2', 'NumberOfGenes']$NumberOfGenes)*100

GeneIsoformRatio_novel[GeneIsoformRatio_novel$Group == 'ccRCC3', 'NumberOfGenes']/sum(GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC3', 'NumberOfGenes']$NumberOfGenes)*100

GeneIsoformRatio_novel[GeneIsoformRatio_novel$Group == 'ccRCC4', 'NumberOfGenes']/sum(GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC4', 'NumberOfGenes']$NumberOfGenes)*100

GeneIsoformRatio_novel[GeneIsoformRatio_novel$Group == 'ccRCC5', 'NumberOfGenes']/sum(GeneIsoformRatio[GeneIsoformRatio$Group == 'ccRCC5', 'NumberOfGenes']$NumberOfGenes)*100
```

### Figure 1I
```{r figure1H}
library(viridis)
color_palette <- viridis(5, option = "D")

# Create bins
isoform_annotations$bins <- cut(isoform_annotations$CellNumber, 
                                         breaks=c(1, 2, 5, 30, 60, 90, 120, 150, Inf), labels=c("1-2","3-5", "6-30", "31-60", "61-90", "91-120", "121-150", ">150"), 
                                         right=FALSE)

# Plot
df_count <- isoform_annotations %>%
  group_by(bins, structural_category, Group) %>%
  summarise(count = n()) %>%
  ungroup()

# Compute the total count for each `bins`
df_total <- df_count %>%
  group_by(bins, Group) %>%
  summarise(total = sum(count)) %>%
  ungroup()

# Calculate the proportion
df_count <- merge(df_count, df_total, by = c("bins", "Group"))
df_count$proportion <- df_count$count / df_count$total

p <- ggplot(df_count, aes(x = bins, y = proportion, fill = structural_category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Proportion", x = "Cell Number Range", title = "The number of cells a transcript is identified") +
  scale_fill_manual(values = color_palette) + 
  theme_classic() +
  facet_wrap(~Group, scales="free_x") +
  theme(legend.position="right") + 
  theme(axis.text.x = element_blank())
  #theme(legend.position="right") + #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  

p
#ggsave("Figure1Ipng", plot = p, bg = "transparent", width = 8, height = 5, dpi=300)
```

### Figure 1J
```{r TranscriptPerGenePerCell}
library(tidyr)

df_normal <- scisoseq_file_normal_filtered2  %>% dplyr::select(BCrev, gene, pbid) %>% dplyr::distinct() %>%
  group_by(BCrev, gene) %>%
  summarize(isoforms = paste(pbid, collapse = ", "),
            .groups = 'drop')

df_ccRCC2 <- scisoseq_file_ccRCC2_filtered2 %>% dplyr::select(BCrev, gene, pbid) %>% dplyr::distinct() %>%
  group_by(BCrev, gene) %>%
  summarize(isoforms = paste(pbid, collapse = ", "),
            .groups = 'drop')

df_ccRCC3 <- scisoseq_file_ccRCC3_filtered2 %>% dplyr::select(BCrev, gene, pbid) %>% dplyr::distinct() %>%
  group_by(BCrev, gene) %>%
  summarize(isoforms = paste(pbid, collapse = ", "),
            .groups = 'drop')

df_ccRCC4 <- scisoseq_file_ccRCC4_filtered2 %>% dplyr::select(BCrev, gene, pbid) %>% dplyr::distinct() %>%
  group_by(BCrev, gene) %>%
  summarize(isoforms = paste(pbid, collapse = ", "),
            .groups = 'drop')

df_ccRCC5 <- scisoseq_file_ccRCC5_filtered2 %>% dplyr::select(BCrev, gene, pbid) %>% dplyr::distinct() %>%
  group_by(BCrev, gene) %>%
  summarize(isoforms = paste(pbid, collapse = ", "),
            .groups = 'drop')
# View the transformed dataframe
df_normal$Group <- 'Normal'
df_ccRCC2$Group <- 'ccRCC2'
df_ccRCC3$Group <- 'ccRCC3'
df_ccRCC4$Group <- 'ccRCC4'
df_ccRCC5$Group <- 'ccRCC5'

df_all <- rbind(df_normal, df_ccRCC2, df_ccRCC3, df_ccRCC4, df_ccRCC5)
```


````{r Figure1J}
# Create bins
# count how many isoforms found in each cell per gene. 
df_all$isoformcount <- sapply(strsplit(df_all$isoforms, ","), length)

df_all$bins <- cut(df_all$isoformcount, breaks=c(0, 2, 4, 7, 10, Inf), labels=c("0-1", "2-3", "4-6", "7-9", ">=10"),  right=FALSE)

df_summary <- df_all %>%
  dplyr::select(gene, bins, Group) %>% dplyr::distinct()

TranscriptPerGenePerCell <- ggplot(df_summary, aes(x = bins)) +
  geom_bar(position = "dodge") +
  labs(y = "Total Number of Genes", x = "Number of Transcripts", title = "Total Number of Genes Having varying Number of Isoforms Per Cell") +
  theme_classic() +
  facet_wrap(~Group, scales = "free_x") +
  theme(legend.position = "bottom") +
theme(axis.text.x = element_blank())

TranscriptPerGenePerCell

#ggsave("Figure1J_transc_gene_cell.png", plot = TranscriptPerGenePerCell, bg = "transparent", width = 6, height = 5, dpi=300)
```

Some genes might have 1 or 2 isoforms across different cells, so they are listed in each category. 


### Supplementary Figure 1C, 1D, and 1E
```{r Supplementary1E}
healthy_counts <- as.data.frame(healthy_isoform_subset[["RNA"]]$counts)
healthy_counts$isoform <- sapply(strsplit(rownames(healthy_isoform_subset), ":"),`[`, 1)
healthy_counts$associated_gene <- sapply(strsplit(rownames(healthy_isoform_subset), ":"),`[`, 2)
healthy_UMI_category <- merge(healthy_counts, isoform_annotations_normal_filtered_w_CellNumber[, c(1,2,6,7,51)], by= c('isoform'))

ccRCC2_counts <- as.data.frame(ccRCC_2_isoform_subset[["RNA"]]$counts)
ccRCC2_counts$isoform <- sapply(strsplit(rownames(ccRCC_2_isoform_subset), ":"),`[`, 1)
ccRCC2_counts$associated_gene <- sapply(strsplit(rownames(ccRCC_2_isoform_subset), ":"),`[`, 2)
ccRCC2_UMI_category <- merge(ccRCC2_counts,isoform_annotations_ccRCC_2_filtered_w_CellNumber[, c(1,2,6,7,51)], by= c('isoform'))


ccRCC3_counts <- as.data.frame(ccRCC_3_isoform_subset[["RNA"]]$counts)
ccRCC3_counts$isoform <- sapply(strsplit(rownames(ccRCC_3_isoform_subset), ":"),`[`, 1)
ccRCC3_counts$associated_gene <- sapply(strsplit(rownames(ccRCC_3_isoform_subset), ":"),`[`, 2)
ccRCC3_UMI_category <- merge(ccRCC3_counts,isoform_annotations_ccRCC_3_filtered_w_CellNumber[, c(1,2,6,7,51)], by= c('isoform'))


ccRCC4_counts <- as.data.frame(ccRCC_4_isoform_subset[["RNA"]]$counts)
ccRCC4_counts$isoform <- sapply(strsplit(rownames(ccRCC_4_isoform_subset), ":"),`[`, 1)
ccRCC4_counts$associated_gene <- sapply(strsplit(rownames(ccRCC_4_isoform_subset), ":"),`[`, 2)
ccRCC4_UMI_category <- merge(ccRCC4_counts,isoform_annotations_ccRCC_4_filtered_w_CellNumber[, c(1,2,6,7,51)], by= c('isoform'))


ccRCC5_counts <- as.data.frame(ccRCC_5_isoform_subset[["RNA"]]$counts)
ccRCC5_counts$isoform <- sapply(strsplit(rownames(ccRCC_5_isoform_subset), ":"),`[`, 1)
ccRCC5_counts$associated_gene <- sapply(strsplit(rownames(ccRCC_5_isoform_subset), ":"),`[`, 2)
ccRCC5_UMI_category <- merge(ccRCC5_counts,isoform_annotations_ccRCC_5_filtered_w_CellNumber[, c(1,2,6,7,51)], by= c('isoform'))
```


```{r analyzeStructuralCategories}
healthy_UMI_category$chrom <- as.character(healthy_UMI_category$chrom)
healthy_UMI_category_long <- pivot_longer(healthy_UMI_category, 
                         cols = ends_with("-1"), 
                        names_to = "CellBarcode", 
                        values_to = "UMICount")

healthy_UMI_category_long$Group <- 'Normal'

ccRCC2_UMI_category$chrom <- as.character(ccRCC2_UMI_category$chrom)
ccRCC2_UMI_category_long <- pivot_longer(ccRCC2_UMI_category, 
                         cols = ends_with("-1"), 
                        names_to = "CellBarcode", 
                        values_to = "UMICount")

ccRCC2_UMI_category_long$Group <- 'ccRCC2'

ccRCC3_UMI_category$chrom <- as.character(ccRCC3_UMI_category$chrom)
ccRCC3_UMI_category_long <- pivot_longer(ccRCC3_UMI_category, 
                         cols = ends_with("-1"), 
                        names_to = "CellBarcode", 
                        values_to = "UMICount")

ccRCC3_UMI_category_long$Group <- 'ccRCC3'


ccRCC4_UMI_category$chrom <- as.character(ccRCC4_UMI_category$chrom)
ccRCC4_UMI_category_long <- pivot_longer(ccRCC4_UMI_category, 
                        cols = ends_with("-1"), 
                        names_to = "CellBarcode", 
                        values_to = "UMICount")

ccRCC4_UMI_category_long$Group <- 'ccRCC4'


ccRCC5_UMI_category$chrom <- as.character(ccRCC5_UMI_category$chrom)
ccRCC5_UMI_category_long <- pivot_longer(ccRCC5_UMI_category, 
                        cols = ends_with("-1"),  
                        names_to = "CellBarcode", 
                        values_to = "UMICount")

ccRCC5_UMI_category_long$Group <- 'ccRCC5'

UMI_category_long <- rbind(healthy_UMI_category_long, ccRCC2_UMI_category_long, ccRCC3_UMI_category_long, ccRCC4_UMI_category_long, ccRCC5_UMI_category_long)

UMI_category_long$structural_category[UMI_category_long$structural_category %in% c('antisense', 'intergenic', 'fusion', 'moreJunctions', 'genic')] <- 'Other'

UMI_category_long_totalUMIperIsoform <- UMI_category_long %>% dplyr::group_by(Group, isoform) %>% dplyr::mutate(totalUMI=sum(UMICount)) %>% dplyr::select(Group, isoform, associated_gene, structural_category, CellNumber, totalUMI) %>% dplyr::distinct()
```


## Supplementary Fig 1E
```{r Supplementary1E}
# distribution across each cell
UMI_category_long_totalUMIperIsoform_plot <- ggplot(UMI_category_long_totalUMIperIsoform, aes(x = structural_category, y = totalUMI)) + 
  geom_boxplot() + 
  #scale_fill_manual(values = color_palette) + 
  labs(title = "Distribution of UMI Counts by Structural Category",
       x = "Structural Category",
       y = "UMI Count") +
  facet_wrap(~Group) +
  theme_classic() +
  scale_y_log10() +
  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position="right") + 
  theme(axis.text.x = element_blank())

#ggsave("Supplementary_Figure1E.png", plot = UMI_category_long_totalUMIperIsoform_plot, bg = "transparent", width = 8, height = 5, dpi=300)
```

saveRDS(UMI_category_long_sum, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/UMI_category_long_sum.RDS')   
saveRDS(UMI_category_long, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/UMI_category_long.RDS')
## Supplementary Fig 1F and 1G
```{r Supplementary1G}
# Create bins
UMI_category_long_totalUMIperIsoform$bins <- cut(UMI_category_long_totalUMIperIsoform$CellNumber, 
                                         breaks=c(0, 5, 30, 60, 90, 120, 150, Inf), labels=c("0-5", "6-30", "31-60", "61-90", "91-120", "121-150", ">150"), 
                                         right=FALSE)
UMI_category_long_totalUMIperIsoform$NormalizedCount <- UMI_category_long_totalUMIperIsoform$totalUMI/UMI_category_long_totalUMIperIsoform$CellNumber

# NIC Isoforms 
UMI_category_long_NIC <- UMI_category_long_totalUMIperIsoform[UMI_category_long_totalUMIperIsoform$structural_category == 'novel_in_catalog',]


Novel_cell_UMI <- ggplot(UMI_category_long_NIC, aes(x = bins, y = NormalizedCount)) + 
  geom_boxplot() + 
  labs(title = "Distribution of UMI Counts of NIC across Cell Range",
       x = "Cell Range",
       y = "UMI Count") +
  facet_wrap(~Group) + theme_classic() +
  scale_y_log10() +
   theme(axis.text.x = element_blank())

Novel_cell_UMI

#ggsave("Supplementary_Figure1D.png", plot = Novel_cell_UMI, bg = "transparent", width = 8, height = 5, dpi=300)

# FSM Isoforms 
UMI_category_long_fsm <- UMI_category_long_totalUMIperIsoform[UMI_category_long_totalUMIperIsoform$structural_category == 'full-splice_match',]

# FSM Isoforms 
fsm_cell_UMI <- ggplot(UMI_category_long_fsm, aes(x = bins, y = NormalizedCount)) + 
  geom_boxplot() + 
  labs(title = "Distribution of UMI Counts of FSM across Cell Range",
       x = "Cell Range",
       y = "UMI Count") +
  facet_wrap(~Group) + theme_classic() +
  scale_y_log10() +
   theme(axis.text.x = element_blank())
fsm_cell_UMI
#ggsave("Supplementary_Figure1C.png", plot = fsm_cell_UMI, bg = "transparent", width = 8, height = 5, dpi=300)
```

# saveDataToDiscuss: saveRDS(UMI_category_long_totalUMIperIsoform, 'Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/Figures/Figures/UMI_category_long_totalUMIperIsoform.RDS')


# Figure 2 : ORFs and Intrinsically Disordered Transcripts
```{r read_ORF_predictions}
ORF_normal <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/BioRxiv/InputFiles/Normal/ORFs/best_ORFs_types.tsv', header = TRUE, sep='\t')
ORF_ccRCC2 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/BioRxiv/InputFiles/ccRCC2/ORFs/best_ORFs_types.tsv', header = TRUE, sep='\t')
ORF_ccRCC3 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/BioRxiv/InputFiles/ccRCC3/ORFs/best_ORFs_types.tsv', header = TRUE, sep='\t')
ORF_ccRCC4 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/BioRxiv/InputFiles/ccRCC4/ORFs/best_ORFs_types.tsv', header = TRUE, sep='\t')
ORF_ccRCC5 <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/BioRxiv/InputFiles/ccRCC5/ORFs/best_ORFs_types.tsv', header = TRUE, sep='\t')
```


## Figure prep
Percentages of ORF Hits across Structures Categories in ccRCC and Normal Samples
```{r calculatePercentages_ccRCCs}
# Merge the data frames
## ccRCC2
ORF_ccRCC2_isoform_annotations_ccRCC2_ORFs <- isoform_annotations_ccRCC_2_filtered_w_CellNumber[isoform_annotations_ccRCC_2_filtered_w_CellNumber$isoform %in% ORF_ccRCC2$PB.id,]

ccRCC2_all <- isoform_annotations_ccRCC_2_filtered_w_CellNumber %>% group_by(structural_category) %>% summarize(count = n())
ccRCC2_orfs <- ORF_ccRCC2_isoform_annotations_ccRCC2_ORFs %>% group_by(structural_category) %>% summarize(count = n())
merged_ccRCC2 <- merge(ccRCC2_all, ccRCC2_orfs, by = "structural_category")

## ccRCC3
ORF_ccRCC3_isoform_annotations_ccRCC3_ORFs <- isoform_annotations_ccRCC_3_filtered_w_CellNumber[isoform_annotations_ccRCC_3_filtered_w_CellNumber$isoform %in% ORF_ccRCC3$PB.id, ]
ccRCC3_all <- isoform_annotations_ccRCC_3_filtered_w_CellNumber %>% group_by(structural_category) %>% summarize(count = n())
ccRCC3_orfs <- ORF_ccRCC3_isoform_annotations_ccRCC3_ORFs %>% group_by(structural_category) %>% summarize(count = n())

merged_ccRCC3 <- merge(ccRCC3_all, ccRCC3_orfs, by = "structural_category")


## ccRCC4
ORF_ccRCC4_isoform_annotations_ccRCC4_ORFs <- isoform_annotations_ccRCC_4_filtered_w_CellNumber[isoform_annotations_ccRCC_4_filtered_w_CellNumber$isoform %in% ORF_ccRCC4$PB.id, ]
ccRCC4_all <- isoform_annotations_ccRCC_4_filtered_w_CellNumber %>% group_by(structural_category) %>% summarize(count = n())
ccRCC4_orfs <- ORF_ccRCC4_isoform_annotations_ccRCC4_ORFs %>% group_by(structural_category) %>% summarize(count = n())
merged_ccRCC4 <- merge(ccRCC4_all, ccRCC4_orfs, by = "structural_category")


## ccRCC5
ORF_ccRCC5_isoform_annotations_ccRCC5_ORFs <- isoform_annotations_ccRCC_5_filtered_w_CellNumber[isoform_annotations_ccRCC_5_filtered_w_CellNumber$isoform %in% ORF_ccRCC5$PB.id, ]
ccRCC5_all <- isoform_annotations_ccRCC_5_filtered_w_CellNumber %>% group_by(structural_category) %>% summarize(count = n())
ccRCC5_orfs <- ORF_ccRCC5_isoform_annotations_ccRCC5_ORFs %>% group_by(structural_category) %>% summarize(count = n())
merged_ccRCC5 <- merge(ccRCC5_all, ccRCC5_orfs, by = "structural_category")

# Normal
ORF_normal_isoform_annotations_normal_ORFs <- isoform_annotations_normal_filtered_w_CellNumber[isoform_annotations_normal_filtered_w_CellNumber$isoform %in% ORF_normal$PB.id, ]

normal_all <- isoform_annotations_normal_filtered_w_CellNumber %>% group_by(structural_category) %>% summarize(count = n())
normal_orfs <- ORF_normal_isoform_annotations_normal_ORFs %>% group_by(structural_category) %>% summarize(count = n())
merged_normal <- merge(normal_all, normal_orfs, by = "structural_category")

merged_ccRCC5$group <- 'ccRCC5'
merged_ccRCC4$group <- 'ccRCC4'
merged_ccRCC3$group <- 'ccRCC3'
merged_ccRCC2$group <- 'ccRCC2'
merged_normal$group <- 'Normal'

medged_ccRCCs <- rbind(merged_ccRCC5, merged_ccRCC4, merged_ccRCC3, merged_ccRCC2, merged_normal)

medged_ccRCCs$percentage <- (medged_ccRCCs$count.y / medged_ccRCCs$count.x) * 100 
medged_ccRCCs_avg <- medged_ccRCCs %>% group_by(structural_category) %>% dplyr::summarise(avg_per = mean(percentage))
# Create the bar graph
ggplot(medged_ccRCCs_avg, aes(x = structural_category, y = avg_per, fill = structural_category)) +
  geom_bar(stat = "identity", position = position_dodge(), color='black') +
#  facet_grid(group ~.) +
  theme_classic() +
  #labs(title = "Percentage Comparison of Structural Categories",
      # x = "Structural Category", y = "Percentage (%)") +
  labs(x = "Structural Category", y = "Percentage (%)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Adjusts the x-axis labels for readability
```



### Figure 2A - manuscript
```{r Figure2A}
library(viridis)
color_palette <- viridis(5, option = "D")
medged_ccRCCs_essentials <- medged_ccRCCs_avg %>% dplyr::filter(structural_category %in% c('full-splice_match', 'incomplete-splice_match', 'novel_in_catalog', 'novel_not_in_catalog'))

# Create the bar graph
ORF_hit <- ggplot(medged_ccRCCs_essentials, aes(x = structural_category, y = avg_per, fill=structural_category)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_text(aes(label=round(avg_per, 2)), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) + 
  labs(x = "Structural Category", y = "Percentage (%)") +
 # labs(title = "Percentage Comparison of Structural Categories",
    #   x = "Structural Category", y = "Percentage (%)") +
    scale_fill_manual(values = color_palette) +  theme_classic() +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Adjusts the x-axis labels for readability 
 theme(axis.text.x = element_blank()) 
ORF_hit
#ggsave("Figure2A.png", plot = ORF_hit, bg = "transparent", width = 8, height = 5, dpi=300)
```

# Figure 2B
```{r Figure2B}
ORF_ccRCC2_seurat <- merge(ORF_ccRCC2_isoform_annotations_ccRCC2_ORFs[,c(1,6,51,52)], ORF_ccRCC2, by.x = 'isoform', by.y = 'PB.id')
ORF_ccRCC3_seurat <- merge(ORF_ccRCC3_isoform_annotations_ccRCC3_ORFs[,c(1,6,51,52)], ORF_ccRCC3, by.x = 'isoform', by.y = 'PB.id')
ORF_ccRCC4_seurat <- merge(ORF_ccRCC4_isoform_annotations_ccRCC4_ORFs[,c(1,6,51,52)], ORF_ccRCC4, by.x = 'isoform', by.y = 'PB.id')
ORF_ccRCC5_seurat <- merge(ORF_ccRCC5_isoform_annotations_ccRCC5_ORFs[,c(1,6,51,52)], ORF_ccRCC5, by.x = 'isoform', by.y = 'PB.id')
ORF_N_seurat <- merge(ORF_normal_isoform_annotations_normal_ORFs[,c(1,6,51,52)], ORF_normal, by.x = 'isoform', by.y = 'PB.id')

ORF_all <- rbind(ORF_N_seurat, ORF_ccRCC2_seurat, ORF_ccRCC3_seurat, ORF_ccRCC4_seurat, ORF_ccRCC5_seurat)

ORFTypes <- ORF_all %>% group_by(Group, structural_category, ORFType) %>% summarise(count = n()) %>% mutate(total = sum(count)) %>%  mutate(proportion = count/total) %>% ungroup()

ORFTypes_necc <- ORFTypes[ORFTypes$structural_category %in% c('full-splice_match', 'incomplete-splice_match', 'novel_in_catalog', 'novel_not_in_catalog'), ]

color_palette2 <- viridis(8, option = "A")

color_palette2 <- brewer.pal(n = 8, name = "Set2")
ORFTypes_graphs <- ggplot(ORFTypes_necc, aes(x = structural_category, y = proportion, fill = ORFType)) +
  geom_bar(stat = "identity", position = "stack") +
  #labs(y = "Percentage", x = "Sample", title = "Proportion of ORF Types") +
  labs(y = "Percentage") +
  scale_fill_manual(values = rev(color_palette2)) + 
  theme_classic() +
  facet_wrap(~Group) +
  #theme(axis.text.x = element_text(angle = 90, hjust = 1))
  theme(axis.text.x = element_blank())
ORFTypes_graphs
#ggsave("Figure2B.png", plot = ORFTypes_graphs, bg = "transparent", width = 8, height = 5)
```

# Figure 2C
```{r Figure2C}
ORF_all_novel <- ORF_all[ORF_all$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),]

ORF_all_novel$bins2 <- cut(ORF_all_novel$CellNumber, 
                                         breaks=c(0, 5, 30, 60, 90, 120, 150, Inf), labels=c("0-5","6-30","31-60", "61-90", "91-120", "121-150", ">150"), 
                                         right=FALSE)
# Plot
df_count_novel <- ORF_all_novel %>%
  group_by(Group, bins2, ORFType) %>%
  summarise(count = n()) %>%
  ungroup()

# Step 2: Compute the total count for each `bins`
df_total_novel <- df_count_novel %>%
  group_by(bins2, Group) %>%
  summarise(total = sum(count)) %>%
  ungroup()

#color_palette3 <- viridis(8, option = "A")
color_palette3 <- brewer.pal(n = 8, name = "Set2")
# Step 3: Calculate the proportion
df_count_novel <- merge(df_count_novel, df_total_novel, by = c("bins2", "Group"))
df_count_novel$proportion <- df_count_novel$count / df_count_novel$total

p_novel <- ggplot(df_count_novel, aes(x = bins2, y = proportion, fill = ORFType)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Proportion", x = "Cell Number Range") +
  #labs(y = "Proportion", x = "Cell Number Range", title = "Distribution of Open Reading Frame (ORF) Types in\nNovel Transcripts as a Function of Cell Prevalence") +
  scale_fill_manual(values = rev(color_palette3)) + 
  theme_classic() +
  facet_wrap(~Group) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
  theme(axis.text.x = element_blank())

#ggsave("Figure2C.png", plot = p_novel, bg = "transparent", width = 8, height = 5)
```

```{r chiSquare}
contingency_table <- table(ORF_all_novel$bins2, ORF_all_novel$ORFType)
test_result <- chisq.test(contingency_table[,c(2,4,6)]) 
test_result_complete <- chisq.test(contingency_table[,c(5,6)]) 
test_result_partial <- chisq.test(contingency_table[,c(1:4)])# excluded internal(-) cases as they are in very limited number
print(test_result) 
print(test_result_complete) 
print(test_result_partial) 
```

## Intrinsically disordered Proteins: IUPRED Results
```{r readiupredresults}
# Normal 
iupred_result_all_Normal <- read.table('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/iupred_all/Normal/output.txt', header = TRUE)
colnames(iupred_result_all_Normal) <- c('isoform', 'Disordered', 'Ordered')

# ccRCC2 
iupred_result_all_ccRCC2 <- read.table('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/iupred_all/ccRCC2/output.txt', header = TRUE)
colnames(iupred_result_all_ccRCC2) <- c('isoform', 'Disordered', 'Ordered')

# ccRCC3 
iupred_result_all_ccRCC3 <- read.table('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/iupred_all/ccRCC3/output.txt', header = TRUE)
colnames(iupred_result_all_ccRCC3) <- c('isoform', 'Disordered', 'Ordered')

# ccRCC4
iupred_result_all_ccRCC4 <- read.table('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/iupred_all/ccRCC4/output.txt', header = TRUE)
colnames(iupred_result_all_ccRCC4) <- c('isoform', 'Disordered', 'Ordered')

# ccRCC5
iupred_result_all_ccRCC5 <- read.table('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/iupred_all/ccRCC5/output.txt', header = TRUE)
colnames(iupred_result_all_ccRCC5) <- c('isoform', 'Disordered', 'Ordered')

```


### Figure 2D
```{r Figure2D}
ORF_ccRCC2_seurat_iupred_FSM <- merge(iupred_result_all_ccRCC2, ORF_ccRCC2_seurat[ORF_ccRCC2_seurat$structural_category  == 'full-splice_match',], by='isoform')
ORF_ccRCC2_seurat_iupred_ISM <- merge(iupred_result_all_ccRCC2, ORF_ccRCC2_seurat[ORF_ccRCC2_seurat$structural_category  == 'incomplete-splice_match',], by='isoform')
ORF_ccRCC2_seurat_iupred_Novel <- merge(iupred_result_all_ccRCC2, ORF_ccRCC2_seurat[ORF_ccRCC2_seurat$structural_category  %in% c('novel_in_catalog', 'novel_not_in_catalog'),], by='isoform')

ORF_ccRCC3_seurat_iupred_FSM <- merge(iupred_result_all_ccRCC3, ORF_ccRCC3_seurat[ORF_ccRCC3_seurat$structural_category  == 'full-splice_match',], by='isoform')
ORF_ccRCC3_seurat_iupred_ISM <- merge(iupred_result_all_ccRCC3, ORF_ccRCC3_seurat[ORF_ccRCC3_seurat$structural_category  == 'incomplete-splice_match',], by='isoform')
ORF_ccRCC3_seurat_iupred_Novel <- merge(iupred_result_all_ccRCC3, ORF_ccRCC3_seurat[ORF_ccRCC3_seurat$structural_category  %in% c('novel_in_catalog', 'novel_not_in_catalog'),], by='isoform')

ORF_ccRCC4_seurat_iupred_FSM <- merge(iupred_result_all_ccRCC4, ORF_ccRCC4_seurat[ORF_ccRCC4_seurat$structural_category  == 'full-splice_match',], by='isoform')
ORF_ccRCC4_seurat_iupred_ISM <- merge(iupred_result_all_ccRCC4, ORF_ccRCC4_seurat[ORF_ccRCC4_seurat$structural_category  == 'incomplete-splice_match',], by='isoform')
ORF_ccRCC4_seurat_iupred_Novel <- merge(iupred_result_all_ccRCC4, ORF_ccRCC4_seurat[ORF_ccRCC4_seurat$structural_category  %in% c('novel_in_catalog', 'novel_not_in_catalog'),], by='isoform')

ORF_ccRCC5_seurat_iupred_FSM <- merge(iupred_result_all_ccRCC5, ORF_ccRCC5_seurat[ORF_ccRCC5_seurat$structural_category  == 'full-splice_match',], by='isoform')
ORF_ccRCC5_seurat_iupred_ISM <- merge(iupred_result_all_ccRCC5, ORF_ccRCC5_seurat[ORF_ccRCC5_seurat$structural_category  == 'incomplete-splice_match',], by='isoform')
ORF_ccRCC5_seurat_iupred_Novel <- merge(iupred_result_all_ccRCC5, ORF_ccRCC5_seurat[ORF_ccRCC5_seurat$structural_category  %in% c('novel_in_catalog', 'novel_not_in_catalog'),], by='isoform')

ORF_seurat_iupred_FSM <- merge(iupred_result_all_Normal, ORF_N_seurat[ORF_N_seurat$structural_category  == 'full-splice_match',], by='isoform')
ORF_seurat_iupred_ISM <- merge(iupred_result_all_Normal, ORF_N_seurat[ORF_N_seurat$structural_category  == 'incomplete-splice_match',], by='isoform')
ORF_seurat_iupred_Novel <- merge(iupred_result_all_Normal, ORF_N_seurat[ORF_N_seurat$structural_category  %in% c('novel_in_catalog', 'novel_not_in_catalog'),], by='isoform')


ORF_seurat_iupred_merge <- rbind(ORF_ccRCC2_seurat_iupred_FSM, ORF_ccRCC2_seurat_iupred_ISM, ORF_ccRCC2_seurat_iupred_Novel, 
                                 ORF_ccRCC3_seurat_iupred_FSM, ORF_ccRCC3_seurat_iupred_ISM, ORF_ccRCC3_seurat_iupred_Novel,
                                 ORF_ccRCC4_seurat_iupred_FSM, ORF_ccRCC4_seurat_iupred_ISM, ORF_ccRCC4_seurat_iupred_Novel,
                                 ORF_ccRCC5_seurat_iupred_FSM, ORF_ccRCC5_seurat_iupred_ISM, ORF_ccRCC5_seurat_iupred_Novel, 
                                 ORF_seurat_iupred_FSM, ORF_seurat_iupred_ISM, ORF_seurat_iupred_Novel)

color_palette <- viridis(5, option = "D")

library(ggpubr)
ORF_seurat_iupred_plot <- ggplot(ORF_seurat_iupred_merge, aes(y=Disordered, x=structural_category, fill=structural_category)) + geom_boxplot(notch=TRUE) + scale_fill_manual(values = color_palette) +  theme_classic() +
  theme(axis.text.x = element_blank()) + stat_compare_means(method = "wilcox.test", label = "p.signif", 
                     comparisons = list(c("full-splice_match", "novel_in_catalog"), 
     c("full-splice_match", "novel_not_in_catalog"), 
     c("incomplete-splice_match", "novel_in_catalog"),
     c("incomplete-splice_match", "full-splice_match"), 
     c("incomplete-splice_match", "novel_not_in_catalog"),
     c("novel_in_catalog", "novel_not_in_catalog")))
  
ORF_seurat_iupred_plot
#ggsave("Figure2D.png", plot = ORF_seurat_iupred_plot, bg = "transparent", width = 8, height = 5)
```
compare_means(
    Disordered ~ structural_category, 
    data = ORF_seurat_iupred_merge, 
    method = "wilcox.test", 
    group.by = "Group", 
    comparisons = list(c("full-splice_match", "novel_in_catalog"), 
     c("full-splice_match", "novel_not_in_catalog"), 
     c("incomplete-splice_match", "novel_in_catalog"),
     c("incomplete-splice_match", "full-splice_match"), 
     c("incomplete-splice_match", "novel_not_in_catalog"))
)


wilcox.test(ORF_seurat_iupred_merge[ORF_seurat_iupred_merge$structural_category == 'full-splice_match', 'Disordered'], ORF_seurat_iupred_merge[ORF_seurat_iupred_merge$structural_category == 'incomplete-splice_match', 'Disordered'])

wilcox.test(ORF_seurat_iupred_merge[ORF_seurat_iupred_merge$structural_category == 'full-splice_match', 'Disordered'], ORF_seurat_iupred_merge[ORF_seurat_iupred_merge$structural_category == 'novel_in_catalog', 'Disordered'])
p-value = 0.009287

### Figure 2E
```{r Figure 2E}
ORF_ccRCC2_seurat_iupred_sub <- merge(ORF_ccRCC2_seurat_iupred_Novel[ORF_ccRCC2_seurat_iupred_Novel$structural_category == 'novel_in_catalog', ], isoform_annotations_ccRCC_2_filtered_w_CellNumber[, c(1,2,7,15)], by=c('isoform'))


ORF_ccRCC2_seurat_iupred_subNNIC <- merge(ORF_ccRCC2_seurat_iupred_Novel[ORF_ccRCC2_seurat_iupred_Novel$structural_category == 'novel_not_in_catalog', ], isoform_annotations_ccRCC_2_filtered_w_CellNumber[, c(1,2,7,15)], by=c('isoform'))

ORF_ccRCC3_seurat_iupred_sub <- merge(ORF_ccRCC3_seurat_iupred_Novel[ORF_ccRCC3_seurat_iupred_Novel$structural_category == 'novel_in_catalog', ], isoform_annotations_ccRCC_3_filtered_w_CellNumber[, c(1,2,7,15)], by=c('isoform'))

ORF_ccRCC3_seurat_iupred_subNNIC <- merge(ORF_ccRCC3_seurat_iupred_Novel[ORF_ccRCC3_seurat_iupred_Novel$structural_category == 'novel_not_in_catalog', ], isoform_annotations_ccRCC_3_filtered_w_CellNumber[, c(1,2,7,15)], by=c('isoform'))

ORF_ccRCC4_seurat_iupred_sub <- merge(ORF_ccRCC4_seurat_iupred_Novel[ORF_ccRCC4_seurat_iupred_Novel$structural_category == 'novel_in_catalog', ], isoform_annotations_ccRCC_4_filtered_w_CellNumber[, c(1,2,7,15)], by=c('isoform'))

ORF_ccRCC4_seurat_iupred_subNNIC <- merge(ORF_ccRCC4_seurat_iupred_Novel[ORF_ccRCC4_seurat_iupred_Novel$structural_category == 'novel_not_in_catalog', ], isoform_annotations_ccRCC_4_filtered_w_CellNumber[, c(1,2,7,15)], by=c('isoform'))

ORF_ccRCC5_seurat_iupred_sub <- merge(ORF_ccRCC5_seurat_iupred_Novel[ORF_ccRCC5_seurat_iupred_Novel$structural_category == 'novel_in_catalog', ], isoform_annotations_ccRCC_5_filtered_w_CellNumber[, c(1,2,7,15)], by=c('isoform'))

ORF_ccRCC5_seurat_iupred_subNNIC <- merge(ORF_ccRCC5_seurat_iupred_Novel[ORF_ccRCC5_seurat_iupred_Novel$structural_category == 'novel_not_in_catalog', ], isoform_annotations_ccRCC_5_filtered_w_CellNumber[, c(1,2,7,15)], by=c('isoform'))

ORF_normal_seurat_iupred_sub <- merge(ORF_seurat_iupred_Novel[ORF_seurat_iupred_Novel$structural_category == 'novel_in_catalog', ], isoform_annotations_normal_filtered_w_CellNumber[, c(1,2,7,15)], by=c('isoform'))

ORF_normal_seurat_iupred_subNNIC <- merge(ORF_seurat_iupred_Novel[ORF_seurat_iupred_Novel$structural_category == 'novel_not_in_catalog', ], isoform_annotations_normal_filtered_w_CellNumber[, c(1,2,7,15)], by=c('isoform'))

ORF_ccRCC5_seurat_iupred_subNNIC$Group <- 'NNC'
ORF_ccRCC4_seurat_iupred_subNNIC$Group <- 'NNC'
ORF_ccRCC3_seurat_iupred_subNNIC$Group <- 'NNC'
ORF_ccRCC2_seurat_iupred_subNNIC$Group <- 'NNC'
ORF_normal_seurat_iupred_subNNIC$Group <- 'NNC'

ORF_ccRCC5_seurat_iupred_sub$Group <- 'NIC'
ORF_ccRCC4_seurat_iupred_sub$Group <- 'NIC'
ORF_ccRCC3_seurat_iupred_sub$Group <- 'NIC'
ORF_ccRCC2_seurat_iupred_sub$Group <- 'NIC'
ORF_normal_seurat_iupred_sub$Group <- 'NIC'


ORF_iupred_subNNIC_NIC <- rbind(ORF_ccRCC5_seurat_iupred_subNNIC,
ORF_ccRCC4_seurat_iupred_subNNIC,
ORF_ccRCC3_seurat_iupred_subNNIC,
ORF_ccRCC2_seurat_iupred_subNNIC,
ORF_normal_seurat_iupred_subNNIC,
ORF_ccRCC5_seurat_iupred_sub,
ORF_ccRCC4_seurat_iupred_sub,
ORF_ccRCC3_seurat_iupred_sub,
ORF_ccRCC2_seurat_iupred_sub,
ORF_normal_seurat_iupred_sub
)


stat_test <- compare_means(
    Disordered ~ subcategory, 
    data = ORF_iupred_subNNIC_NIC, 
    method = "wilcox.test", 
    group.by = "Group", 
    comparisons = list(c('combination_of_known_splicesites', 'intron_retention'),
        c("combination_of_known_splicesites", "combination_of_known_junctions"), 
        c("combination_of_known_junctions", "intron_retention"),
        c("at_least_one_novel_splicesite", "intron_retention")
    )
)

stat_test <- stat_test %>%
    mutate(y.position = c(110, 130, 150, 110))

figure2e <- ggplot(ORF_iupred_subNNIC_NIC, aes(y = Disordered, x = subcategory, fill = Group)) + 
    geom_boxplot(notch = TRUE) + 
    facet_wrap(~Group, scales = 'free_x') + 
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 1)) + 
    scale_fill_manual(values = color_palette[3:4]) +  
    theme_classic() + 
    theme(axis.text.x = element_blank()) +
    stat_pvalue_manual(stat_test, label = "p.signif", tip.length = 0.02)
#ggsave("Figure2E.png", plot = figure2e, bg = "transparent", width = 8, height = 5)
```

## Supplementary 2B
## Additional Test to check 
```{r filterFSM_RM}
# Filter only FSM reference match transcripts
ORF_ccRCC2_seurat_forRM <- merge(ORF_ccRCC2_isoform_annotations_ccRCC2_ORFs[,c(1,6,7, 15, 51,52)], ORF_ccRCC2, by.x = 'isoform', by.y = 'PB.id')
ORF_ccRCC3_seurat_forRM <- merge(ORF_ccRCC3_isoform_annotations_ccRCC3_ORFs[,c(1,6,7, 15, 51,52)], ORF_ccRCC3, by.x = 'isoform', by.y = 'PB.id')
ORF_ccRCC4_seurat_forRM <- merge(ORF_ccRCC4_isoform_annotations_ccRCC4_ORFs[,c(1,6,7,15, 51,52)], ORF_ccRCC4, by.x = 'isoform', by.y = 'PB.id')
ORF_ccRCC5_seurat_forRM <- merge(ORF_ccRCC5_isoform_annotations_ccRCC5_ORFs[,c(1,6,7,15, 51,52)], ORF_ccRCC5, by.x = 'isoform', by.y = 'PB.id')
ORF_N_seurat_forRM <- merge(ORF_normal_isoform_annotations_normal_ORFs[,c(1,6,7, 15, 51,52)], ORF_normal, by.x = 'isoform', by.y = 'PB.id')


### Figure 2D - Reference match
ORF_ccRCC2_seurat_iupred_FSM_forRM  <- merge(iupred_result_FSM_ccRCC2, ORF_ccRCC2_seurat_forRM , by='isoform')
ORF_ccRCC3_seurat_iupred_FSM_forRM  <- merge(iupred_result_FSM_ccRCC3, ORF_ccRCC3_seurat_forRM, by='isoform')
ORF_ccRCC4_seurat_iupred_FSM_forRM  <- merge(iupred_result_FSM_ccRCC4, ORF_ccRCC4_seurat_forRM, by='isoform')
ORF_ccRCC5_seurat_iupred_FSM_forRM  <- merge(iupred_result_FSM_ccRCC5, ORF_ccRCC5_seurat_forRM, by='isoform')
ORF_seurat_iupred_FSM_forRM  <- merge(iupred_result_FSM, ORF_N_seurat_forRM, by='isoform')
ORF_seurat_iupred_merge_forRM  <- rbind(ORF_ccRCC2_seurat_iupred_FSM_forRM,
                                        ORF_ccRCC3_seurat_iupred_FSM_forRM,
                                 ORF_ccRCC4_seurat_iupred_FSM_forRM,
                                 ORF_ccRCC5_seurat_iupred_FSM_forRM,
                                 ORF_seurat_iupred_FSM_forRM)

color_palette <- viridis(5, option = "D")

ORF_seurat_iupred_plot_forRM  <- ggplot(ORF_seurat_iupred_merge_forRM , aes(y=Disordered, x=subcategory, fill=subcategory)) + geom_boxplot(notch=TRUE) + scale_fill_manual(values = color_palette) +  theme_classic() +
  theme(axis.text.x = element_blank()) + stat_compare_means(method = "wilcox.test", label = "p.signif", 
                     comparisons = list(c("alternative_3end", "alternative_5end"), 
     c("alternative_3end", "reference_match"), 
     c("reference_match", "alternative_5end"),
     c("reference_match", "mono-exon"), 
     c("alternative_3end", "mono-exon"),
      c("alternative_5end", "mono-exon"),
     c("alternative_5end", "alternative_3end5end"),
     c("alternative_3end", "alternative_3end5end"),
     c("reference_match", "alternative_3end5end"),
     c("reference_match", "mono-exon")))
# Figure Supplementary 2B
ORF_seurat_iupred_plot_forRM 
#ggsave("Figure2D.png", plot = ORF_seurat_iupred_plot, bg = "transparent", width = 8, height = 5)
```

## Supplementary 2C
compare reference match and novel only
```{r RFM_Novel}
ORF_seurat_iupred_FSM_forRM_onlyRM <- ORF_seurat_iupred_FSM_forRM %>% dplyr::filter(subcategory == 'reference_match')
ORF_seurat_iupred_FSM_forRM_RM_novel <- rbind(ORF_seurat_iupred_FSM_forRM_onlyRM[, c(1,2,3, 4, 5)], ORF_iupred_subNNIC_NIC[c(1,2,3,4,10)])

paired_data <- ORF_seurat_iupred_FSM_forRM_RM_novel %>%
  select(isoform, Disordered, structural_category, associated_gene.x)

library(tidyr)
long_data <- paired_data %>%
  pivot_longer(cols = c("Disordered"), 
               names_to = "Condition", 
               values_to = "Value")

ggplot(long_data, aes(x = structural_category, y = Value, color = structural_category)) +
  geom_boxplot() +  # Boxplot without outliers
  #geom_jitter(width = 0.2) +  # Add jittered points
  #geom_line(aes(group = associated_gene.x), color = "gray", size = 0.4) +  # Connect dots from the same gene
  scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07")) +  # Optional palette
  #stat_compare_means(paired = TRUE) +  # Add paired statistical comparison
  theme_classic() +  # Clean theme
  labs(x = "Structural Category", y = "Value", color = "Structural Category") +  stat_compare_means(method = "wilcox.test", label = "p.signif", 
                     comparisons = list(c("novel_in_catalog", "novel_not_in_catalog"), 
     c("full-splice_match", "novel_in_catalog"), 
     c("full-splice_match", "novel_not_in_catalog")))
```

## Supplementary 2A
```{r calculatePercentages_ccRCCs_Sub}
# Merge the data frames
## ccRCC2
isoform_annotations_ccRCC_2_filtered_w_CellNumber$structural_category[isoform_annotations_ccRCC_2_filtered_w_CellNumber$structural_category %in% c('antisense', 'intergenic', 'fusion', 'moreJunctions', 'genic')] <- 'Other'
isoform_annotations_ccRCC_3_filtered_w_CellNumber$structural_category[isoform_annotations_ccRCC_3_filtered_w_CellNumber$structural_category %in% c('antisense', 'intergenic', 'fusion', 'moreJunctions', 'genic')] <- 'Other'
isoform_annotations_ccRCC_4_filtered_w_CellNumber$structural_category[isoform_annotations_ccRCC_4_filtered_w_CellNumber$structural_category %in% c('antisense', 'intergenic', 'fusion', 'moreJunctions', 'genic')] <- 'Other'
isoform_annotations_ccRCC_5_filtered_w_CellNumber$structural_category[isoform_annotations_ccRCC_5_filtered_w_CellNumber$structural_category %in% c('antisense', 'intergenic', 'fusion', 'moreJunctions', 'genic')] <- 'Other'
isoform_annotations_normal_filtered_w_CellNumber$structural_category[isoform_annotations_normal_filtered_w_CellNumber$structural_category %in% c('antisense', 'intergenic', 'fusion', 'moreJunctions', 'genic')] <- 'Other'

ccRCC2_all_sub <- isoform_annotations_ccRCC_2_filtered_w_CellNumber %>% group_by(structural_category, subcategory) %>% summarize(count = n())
ccRCC2_orfs_sub <- ORF_ccRCC2_isoform_annotations_ccRCC2_ORFs %>% group_by(structural_category, subcategory) %>% summarize(count = n())
merged_ccRCC2_sub <- merge(ccRCC2_all_sub, ccRCC2_orfs_sub, by = c("structural_category", "subcategory"))

## ccRCC3
ccRCC3_all_sub <- isoform_annotations_ccRCC_3_filtered_w_CellNumber %>% group_by(structural_category, subcategory) %>% summarize(count = n())
ccRCC3_orfs_sub <- ORF_ccRCC3_isoform_annotations_ccRCC3_ORFs %>% group_by(structural_category, subcategory) %>% summarize(count = n())

merged_ccRCC3_sub <- merge(ccRCC3_all_sub, ccRCC3_orfs_sub, by = c("structural_category", "subcategory"))


## ccRCC4
ccRCC4_all_sub <- isoform_annotations_ccRCC_4_filtered_w_CellNumber %>% group_by(structural_category, subcategory) %>% summarize(count = n())
ccRCC4_orfs_sub <- ORF_ccRCC4_isoform_annotations_ccRCC4_ORFs %>% group_by(structural_category, subcategory) %>% summarize(count = n())
merged_ccRCC4_sub <- merge(ccRCC4_all_sub, ccRCC4_orfs_sub, by = c("structural_category", "subcategory"))


## ccRCC5
ccRCC5_all_sub <- isoform_annotations_ccRCC_5_filtered_w_CellNumber %>% group_by(structural_category, subcategory) %>% summarize(count = n())
ccRCC5_orfs_sub <- ORF_ccRCC5_isoform_annotations_ccRCC5_ORFs %>% group_by(structural_category, subcategory) %>% summarize(count = n())
merged_ccRCC5_sub <- merge(ccRCC5_all_sub, ccRCC5_orfs_sub, by = c("structural_category", "subcategory"))

# Normal
normal_all_sub <- isoform_annotations_normal_filtered_w_CellNumber %>% group_by(structural_category, subcategory) %>% summarize(count = n())
normal_orfs_sub <- ORF_normal_isoform_annotations_normal_ORFs %>% group_by(structural_category, subcategory) %>% summarize(count = n())
merged_normal_sub <- merge(normal_all_sub, normal_orfs_sub, by = c("structural_category", "subcategory"))

merged_ccRCC5_sub$group <- 'ccRCC5'
merged_ccRCC4_sub$group <- 'ccRCC4'
merged_ccRCC3_sub$group <- 'ccRCC3'
merged_ccRCC2_sub$group <- 'ccRCC2'
merged_normal_sub$group <- 'Normal'

medged_ccRCCs_sub <- rbind(merged_ccRCC5_sub, merged_ccRCC4_sub, merged_ccRCC3_sub, merged_ccRCC2_sub, merged_normal_sub)

medged_ccRCCs_sub$percentage <- (medged_ccRCCs_sub$count.y / medged_ccRCCs_sub$count.x) * 100 
medged_ccRCCs_avg_sub <- medged_ccRCCs_sub %>% group_by(structural_category, subcategory) %>% dplyr::summarise(avg_per = mean(percentage))
# Create the bar graph
# Figure Supplementary 2A
ggplot(medged_ccRCCs_avg_sub, aes(x = subcategory, y = avg_per)) +
    geom_bar(stat = "identity", position = position_dodge(), color='black') +
    facet_wrap(~structural_category, scales="free_x", nrow=1) +
    geom_text(aes(label=round(avg_per, 2)), vjust=1.6, color="white",
              position = position_dodge(0.9), size=3.5) + 
theme_classic() +
    labs(x = "Isoform Subcategories", y = "Percentage (%)") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


isoform_annotations_ccRCC_2_filtered_w_CellNumber[isoform_annotations_ccRCC_2_filtered_w_CellNumber$isoform %in% ORF_ccRCC2[ORF_ccRCC2$ORFType %in% c('complete (+)', 'complete (-)'), 'PB.id'] & isoform_annotations_ccRCC_2_filtered_w_CellNumber$associated_gene.x == 'NNMT' & isoform_annotations_ccRCC_2_filtered_w_CellNumber$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),] # 8 NNMT novel isoform 

isoform_annotations_ccRCC_5_filtered_w_CellNumber[isoform_annotations_ccRCC_5_filtered_w_CellNumber$isoform %in% ORF_ccRCC5[ORF_ccRCC5$ORFType %in% c('complete (+)', 'complete (-)'), 'PB.id'] & isoform_annotations_ccRCC_5_filtered_w_CellNumber$associated_gene.x == 'NNMT' & isoform_annotations_ccRCC_5_filtered_w_CellNumber$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'),]

