---
title: "Manuscript_4"
author: "Tülay Karakulak"
date: "2024-09-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Codes to generate Figure 4A-C and Figure Supplementary 3
```{r libraries}
library(dplyr)
library(ggplot2)
library(knitr)
library(rmarkdown)
library(harmony)
library(Matrix)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(ggpubr)
theme_set(theme_cowplot())
set.seed(12345)
```


```{r ReadIds}
IsoMatch_All <- read.csv('/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/TamaIDs/PbIds_TamaIds.tsv', sep=' ', header=FALSE)
colnames(IsoMatch_All) <- c('TamaID', 'Sample', 'PBid')
head(IsoMatch_All)
```


```{r mergeAnnotations}
isoform_annotations_normal_filtered <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_normal_filtered_w_CellNumber.RDS')
isoform_annotations_ccRCC_2_filtered <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_ccRCC_2_filtered_w_CellNumber.RDS')
isoform_annotations_ccRCC_3_filtered <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_ccRCC_3_filtered_w_CellNumber.RDS')
isoform_annotations_ccRCC_4_filtered <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_ccRCC_4_filtered_w_CellNumber.RDS')
isoform_annotations_ccRCC_5_filtered <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotation#isoform_annotations_ccRCC_5_filtered_w_CellNumber.RDS')

isoform_annotations_normal_tama <- merge(isoform_annotations_normal_filtered, IsoMatch_All[IsoMatch_All$Sample == 'Normal',], by.x = 'isoform', by.y='PBid')
isoform_annotations_ccRCC2_tama <- merge(isoform_annotations_ccRCC_2_filtered, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC2',], by.x = 'isoform', by.y='PBid')
isoform_annotations_ccRCC3_tama <- merge(isoform_annotations_ccRCC_3_filtered, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC3',], by.x = 'isoform', by.y='PBid')
isoform_annotations_ccRCC4_tama <- merge(isoform_annotations_ccRCC_4_filtered, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC4',], by.x = 'isoform', by.y='PBid')
isoform_annotations_ccRCC5_tama <- merge(isoform_annotations_ccRCC_5_filtered, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC5',], by.x = 'isoform', by.y='PBid')
```



```{r NumberOfMatchingIsoforms2}
library(UpSetR)
IsoformIds_list_samples2 <- list(Normal = unique(isoform_annotations_normal_tama$TamaID), ccRCC2 = unique(isoform_annotations_ccRCC2_tama$TamaID), ccRCC3 = unique(isoform_annotations_ccRCC3_tama$TamaID), ccRCC4 = unique(isoform_annotations_ccRCC4_tama$TamaID), ccRCC5 = unique(isoform_annotations_ccRCC5_tama$TamaID))

IsoformMatching2 <- upset(fromList(IsoformIds_list_samples2), order.by = 'freq', nsets = 6, number.angles = 45, point.size = 3.5, line.size = 2, mainbar.y.label = "# Isoforms", sets.x.label = "Number of Isoforms", text.scale = 2)

IsoformMatching2

library(ggVennDiagram)

ggVennDiagram(IsoformIds_list_samples2) + scale_fill_gradient(low = "grey", high='darkred')

png("Figure4A.png", width = 20 * 300, height = 10 * 300, res = 300, bg = "transparent")
print(IsoformMatching2)
dev.off()
```


```{r readIsoforms}
ccRCC_Marker_Annotations <- read.table( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/ccRCC_CellType_Revised.tsv', sep='\t', header = TRUE)
ccRCC_Marker_Annotations$BC <- sapply(strsplit(ccRCC_Marker_Annotations$CellBarcode, "-"), function(x) x[1])
head(ccRCC_Marker_Annotations)
```


length(unique(c(seurat_annotations_normal_tama$TamaID, seurat_annotations_ccRCC_2_tama$TamaID, seurat_annotations_ccRCC_4_tama$TamaID, seurat_annotations_ccRCC_5_tama$TamaID, seurat_annotations_ccRCC_3_tama$TamaID)))
[1] 311586

length(unique(c(seurat_annotations_normal_tama[seurat_annotations_normal_tama$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'], seurat_annotations_ccRCC_2_tama[seurat_annotations_ccRCC_2_tama$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'], seurat_annotations_ccRCC_4_tama[seurat_annotations_ccRCC_4_tama$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'], seurat_annotations_ccRCC_5_tama[seurat_annotations_ccRCC_5_tama$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'], seurat_annotations_ccRCC_3_tama[seurat_annotations_ccRCC_3_tama$structural_category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'])))
[1] 83009

%26 are novel

# Supplementary Fig 3
```{r JaccardSimilarity}
jaccard_similarity_Novel_filtered <- function(isoAnnotation1, isoAnnotation2) {
   
    #isoAnnotation1 <- isoAnnotation1[isoAnnotation1$number_of_Cellbarcodes >= 3, ]
    #isoAnnotation2 <- isoAnnotation2[isoAnnotation2$number_of_Cellbarcodes >= 3, ]
  
    intersect1_2 <-length(intersect(isoAnnotation1[isoAnnotation1$structural_category %in% c('novel_not_in_catalog', 'novel_in_catalog'), 'TamaID'],isoAnnotation2[isoAnnotation2$structural_category %in% c('novel_not_in_catalog', 'novel_in_catalog'), 'TamaID']))
                        
  distinct1 <- length(isoAnnotation1[isoAnnotation1$structural_category %in% c('novel_not_in_catalog', 'novel_in_catalog'), 'TamaID']) - intersect1_2
  distinct2 <- length(isoAnnotation2[isoAnnotation2$structural_category %in% c('novel_not_in_catalog', 'novel_in_catalog'), 'TamaID']) - intersect1_2
  union1_2 <- intersect1_2 + distinct1 + distinct2
  similarity <- intersect1_2/union1_2
  return(similarity)
}

JC1_1 <- 1
JC1_2 <- jaccard_similarity_Novel_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC2_tama)
JC1_3 <- jaccard_similarity_Novel_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC3_tama)
JC1_4 <- jaccard_similarity_Novel_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC4_tama)
JC1_5 <- jaccard_similarity_Novel_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC5_tama)

JC2_2 <- 1
JC2_3 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC3_tama)
JC2_4 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC4_tama)
JC2_5 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC5_tama)

JC3_3 <- 1
JC3_4 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC4_tama)
JC3_5 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC5_tama)
JC4_4 <- 1
JC4_5 <- jaccard_similarity_Novel_filtered(isoform_annotations_ccRCC4_tama, isoform_annotations_ccRCC5_tama)
JC5_5 <- 1

jaccard_matrix <- matrix(0, nrow = 5, ncol = 5)
diag(jaccard_matrix) <- 1
jaccard_matrix[lower.tri(jaccard_matrix)] <- c(JC1_2, JC1_3, JC1_4, JC1_5, JC2_3, JC2_4, JC2_5, JC3_4, JC3_5, JC4_5)
jaccard_matrix[upper.tri(jaccard_matrix)] <- t(jaccard_matrix)[upper.tri(jaccard_matrix)]

# Generate the heatmap
library(pheatmap)
colnames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
rownames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
breaks <- seq(0, 0.4, length.out = 100)
colors <- colorRampPalette(c("blue", "red"))(99)

FigureSupp5A <- pheatmap(jaccard_matrix, 
         color = colors, 
         breaks = breaks,
         clustering_distance_rows = "euclidean", 
         clustering_distance_cols = "euclidean",
         display_numbers = TRUE, 
         number_color = "white", 
         cluster_rows = FALSE,
         cluster_cols = FALSE)

ggsave("Supplementary_Figure_7A.png", plot = FigureSupp5A, bg = "transparent", width = 5, height = 4)
```

```{r 3Allcellssimilarity}
jaccard_similarity_filtered <- function(isoAnnotation1, isoAnnotation2) {
  
    intersect1_2 <-length(intersect(isoAnnotation1$TamaID,isoAnnotation2$TamaID))
                        
  distinct1 <- length(isoAnnotation1$TamaID) - intersect1_2
  distinct2 <- length(isoAnnotation2$TamaID) - intersect1_2
  union1_2 <- intersect1_2 + distinct1 + distinct2
  similarity <- intersect1_2/union1_2
  return(similarity)
}


JC1_1 <- 1
JC1_2 <- jaccard_similarity_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC2_tama)
JC1_3 <- jaccard_similarity_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC3_tama)
JC1_4 <- jaccard_similarity_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC4_tama)
JC1_5 <- jaccard_similarity_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC5_tama)

JC2_2 <- 1
JC2_3 <- jaccard_similarity_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC3_tama)
JC2_4 <- jaccard_similarity_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC4_tama)
JC2_5 <- jaccard_similarity_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC5_tama)

JC3_3 <- 1
JC3_4 <- jaccard_similarity_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC4_tama)
JC3_5 <- jaccard_similarity_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC5_tama)
JC4_4 <- 1
JC4_5 <- jaccard_similarity_filtered(isoform_annotations_ccRCC4_tama, isoform_annotations_ccRCC5_tama)
JC5_5 <- 1

jaccard_matrix <- matrix(0, nrow = 5, ncol = 5)
diag(jaccard_matrix) <- 1
jaccard_matrix[lower.tri(jaccard_matrix)] <- c(JC1_2, JC1_3, JC1_4, JC1_5, JC2_3, JC2_4, JC2_5, JC3_4, JC3_5, JC4_5)
jaccard_matrix[upper.tri(jaccard_matrix)] <- t(jaccard_matrix)[upper.tri(jaccard_matrix)]

# Generate the heatmap
library(pheatmap)
colnames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
rownames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
breaks <- seq(0, 0.4, length.out = 100)
colors <- colorRampPalette(c("blue", "red"))(99)

figure5A_ALL <- pheatmap(jaccard_matrix, 
         color = colors, 
         breaks = breaks,
         clustering_distance_rows = "euclidean", 
         clustering_distance_cols = "euclidean",
         display_numbers = TRUE, 
         number_color = "white", 
         cluster_rows = FALSE,
         cluster_cols = FALSE)

ggsave("Supplementary_Figure_7A_all.png", plot = figure5A_ALL, bg = "transparent", width = 5, height = 4)

```



```{r 3cellssimilarityFSM}
jaccard_similarity_FSM_filtered <-  function(isoAnnotation1, isoAnnotation2) {
   
  
    intersect1_2 <-length(intersect(isoAnnotation1[isoAnnotation1$structural_category %in% c('full-splice_match'), 'TamaID'],isoAnnotation2[isoAnnotation2$structural_category %in% c('full-splice_match'), 'TamaID']))
                        
  distinct1 <- length(isoAnnotation1[isoAnnotation1$structural_category %in% c('full-splice_match'), 'TamaID']) - intersect1_2
  distinct2 <- length(isoAnnotation2[isoAnnotation2$structural_category %in% c('full-splice_match'), 'TamaID']) - intersect1_2
  union1_2 <- intersect1_2 + distinct1 + distinct2
  similarity <- intersect1_2/union1_2
  return(similarity)
}


JC1_1 <- 1
JC1_2 <- jaccard_similarity_FSM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC2_tama)
JC1_3 <- jaccard_similarity_FSM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC3_tama)
JC1_4 <- jaccard_similarity_FSM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC4_tama)
JC1_5 <- jaccard_similarity_FSM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC5_tama)

JC2_2 <- 1
JC2_3 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC3_tama)
JC2_4 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC4_tama)
JC2_5 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC5_tama)

JC3_3 <- 1
JC3_4 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC4_tama)
JC3_5 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC5_tama)
JC4_4 <- 1
JC4_5 <- jaccard_similarity_FSM_filtered(isoform_annotations_ccRCC4_tama, isoform_annotations_ccRCC5_tama)
JC5_5 <- 1

jaccard_matrix <- matrix(0, nrow = 5, ncol = 5)
diag(jaccard_matrix) <- 1
jaccard_matrix[lower.tri(jaccard_matrix)] <- c(JC1_2, JC1_3, JC1_4, JC1_5, JC2_3, JC2_4, JC2_5, JC3_4, JC3_5, JC4_5)
jaccard_matrix[upper.tri(jaccard_matrix)] <- t(jaccard_matrix)[upper.tri(jaccard_matrix)]

# Generate the heatmap
library(pheatmap)
colnames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
rownames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
breaks <- seq(0, 0.4, length.out = 100)
colors <- colorRampPalette(c("blue", "red"))(99)

figure5a_fsm <- pheatmap(jaccard_matrix, 
         color = colors, 
         breaks = breaks,
         clustering_distance_rows = "euclidean", 
         clustering_distance_cols = "euclidean",
         display_numbers = TRUE, 
         number_color = "white", 
         cluster_rows = FALSE,
         cluster_cols = FALSE)


ggsave("Supplementary_Figure_7A_fsm.png", plot = figure5a_fsm, bg = "transparent", width = 5, height = 4)
```


```{r 3cellssimilarityISM}
jaccard_similarity_ISM_filtered <-  function(isoAnnotation1, isoAnnotation2) {

    intersect1_2 <-length(intersect(isoAnnotation1[isoAnnotation1$structural_category %in% c('incomplete-splice_match'), 'TamaID'],isoAnnotation2[isoAnnotation2$structural_category %in% c('incomplete-splice_match'), 'TamaID']))
                        
  distinct1 <- length(isoAnnotation1[isoAnnotation1$structural_category %in% c('incomplete-splice_match'), 'TamaID']) - intersect1_2
  distinct2 <- length(isoAnnotation2[isoAnnotation2$structural_category %in% c('incomplete-splice_match'), 'TamaID']) - intersect1_2
  union1_2 <- intersect1_2 + distinct1 + distinct2
  similarity <- intersect1_2/union1_2
  return(similarity)
}

  
JC1_1 <- 1
JC1_2 <- jaccard_similarity_ISM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC2_tama)
JC1_3 <- jaccard_similarity_ISM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC3_tama)
JC1_4 <- jaccard_similarity_ISM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC4_tama)
JC1_5 <- jaccard_similarity_ISM_filtered(isoform_annotations_normal_tama, isoform_annotations_ccRCC5_tama)

JC2_2 <- 1
JC2_3 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC3_tama)
JC2_4 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC4_tama)
JC2_5 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC2_tama, isoform_annotations_ccRCC5_tama)

JC3_3 <- 1
JC3_4 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC4_tama)
JC3_5 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC5_tama)
JC4_4 <- 1
JC4_5 <- jaccard_similarity_ISM_filtered(isoform_annotations_ccRCC4_tama, isoform_annotations_ccRCC5_tama)
JC5_5 <- 1

jaccard_matrix <- matrix(0, nrow = 5, ncol = 5)
diag(jaccard_matrix) <- 1
jaccard_matrix[lower.tri(jaccard_matrix)] <- c(JC1_2, JC1_3, JC1_4, JC1_5, JC2_3, JC2_4, JC2_5, JC3_4, JC3_5, JC4_5)
jaccard_matrix[upper.tri(jaccard_matrix)] <- t(jaccard_matrix)[upper.tri(jaccard_matrix)]

# Generate the heatmap
library(pheatmap)
colnames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
rownames(jaccard_matrix) <- c("Normal", "ccRCC2", "ccRCC3", "ccRCC4", "ccRCC5")
breaks <- seq(0, 0.4, length.out = 100)
colors <- colorRampPalette(c("blue", "red"))(99)

figure5a_ism <- pheatmap(jaccard_matrix, 
         color = colors, 
         breaks = breaks,
         clustering_distance_rows = "euclidean", 
         clustering_distance_cols = "euclidean",
         display_numbers = TRUE, 
         number_color = "white", 
         cluster_rows = FALSE,
         cluster_cols = FALSE)



ggsave("Supplementary_Figure_7A_ism.png", plot = figure5a_ism, bg = "transparent", width = 5, height = 4)
```

# Figure 4B
```{r isoformmatch}
AllMatchingIDs <- intersect(intersect(intersect(intersect(isoform_annotations_ccRCC2_tama$TamaID, isoform_annotations_ccRCC3_tama$TamaID), isoform_annotations_ccRCC4_tama$TamaID), isoform_annotations_ccRCC5_tama$TamaID), isoform_annotations_normal_tama$TamaID)
```

```{r readIsoMatches}
CellNumberInNormal <- as.data.frame(isoform_annotations_normal_tama[isoform_annotations_normal_tama$TamaID %in% AllMatchingIDs, 'CellNumber'])
CellNumberInccRCC2 <- as.data.frame(isoform_annotations_ccRCC2_tama[isoform_annotations_ccRCC2_tama$TamaID %in% AllMatchingIDs, 'CellNumber'])
CellNumberInccRCC3 <- as.data.frame(isoform_annotations_ccRCC3_tama[isoform_annotations_ccRCC3_tama$TamaID %in% AllMatchingIDs, 'CellNumber'])
CellNumberInccRCC4 <- as.data.frame(isoform_annotations_ccRCC4_tama[isoform_annotations_ccRCC4_tama$TamaID %in% AllMatchingIDs, 'CellNumber'])
CellNumberInccRCC5 <- as.data.frame(isoform_annotations_ccRCC5_tama[isoform_annotations_ccRCC5_tama$TamaID %in% AllMatchingIDs, 'CellNumber'])

CellNumberInNormal_nonmatch <- as.data.frame(isoform_annotations_normal_tama[!isoform_annotations_normal_tama$TamaID %in% AllMatchingIDs, 'CellNumber'])
CellNumberInccRCC2_nonmatch <- as.data.frame(isoform_annotations_ccRCC2_tama[!isoform_annotations_ccRCC2_tama$TamaID %in% AllMatchingIDs, 'CellNumber'])
CellNumberInccRCC3_nonmatch <- as.data.frame(isoform_annotations_ccRCC3_tama[!isoform_annotations_ccRCC3_tama$TamaID %in% AllMatchingIDs, 'CellNumber'])
CellNumberInccRCC4_nonmatch <- as.data.frame(isoform_annotations_ccRCC4_tama[!isoform_annotations_ccRCC4_tama$TamaID %in% AllMatchingIDs, 'CellNumber'])
CellNumberInccRCC5_nonmatch <- as.data.frame(isoform_annotations_ccRCC5_tama[!isoform_annotations_ccRCC5_tama$TamaID %in% AllMatchingIDs, 'CellNumber'])



CellNumberInNormal$Group <- 'Normal'
CellNumberInccRCC2$Group <- 'ccRCC2'
CellNumberInccRCC3$Group <- 'ccRCC3'
CellNumberInccRCC4$Group <- 'ccRCC4'
CellNumberInccRCC5$Group <- 'ccRCC5'

CellNumberInNormal$Status <- 'Matched'
CellNumberInccRCC2$Status <- 'Matched'
CellNumberInccRCC3$Status <- 'Matched'
CellNumberInccRCC4$Status <- 'Matched'
CellNumberInccRCC5$Status <- 'Matched'


CellNumberInNormal_nonmatch$Group <- 'Normal'
CellNumberInccRCC2_nonmatch$Group <- 'ccRCC2'
CellNumberInccRCC3_nonmatch$Group <- 'ccRCC3'
CellNumberInccRCC4_nonmatch$Group <- 'ccRCC4'
CellNumberInccRCC5_nonmatch$Group <- 'ccRCC5'

CellNumberInNormal_nonmatch$Status <- 'UnMatched'
CellNumberInccRCC2_nonmatch$Status <- 'UnMatched'
CellNumberInccRCC3_nonmatch$Status <- 'UnMatched'
CellNumberInccRCC4_nonmatch$Status <- 'UnMatched'
CellNumberInccRCC5_nonmatch$Status <- 'UnMatched'

colnames(CellNumberInccRCC5) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC4) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC3) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC2) <- c('CellNumber', 'Group', 'Status')


colnames(CellNumberInNormal) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC4_nonmatch) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC3_nonmatch) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInccRCC2_nonmatch) <- c('CellNumber', 'Group', 'Status')
colnames(CellNumberInNormal_nonmatch) <- c('CellNumber', 'Group', 'Status')

colnames(CellNumberInccRCC5_nonmatch) <- c('CellNumber', 'Group', 'Status')
All_Matched_cellNumber <- rbind(CellNumberInNormal, CellNumberInccRCC2, CellNumberInccRCC3, CellNumberInccRCC4, CellNumberInccRCC5)
colnames(All_Matched_cellNumber) <- c('CellNumber', 'Group', 'Status')
All_UnMatched_cellNumber <-  rbind(CellNumberInNormal_nonmatch, CellNumberInccRCC2_nonmatch, CellNumberInccRCC4_nonmatch,CellNumberInccRCC3_nonmatch, CellNumberInccRCC5_nonmatch)
colnames(All_UnMatched_cellNumber) <- c('CellNumber', 'Group', 'Status')
All_Matched_UnMatched <- rbind(All_Matched_cellNumber, All_UnMatched_cellNumber)
head(All_Matched_UnMatched)
```


```{r DensityPlot}
figure5bsupp <- ggplot(All_Matched_UnMatched, aes(CellNumber, fill=Status)) + geom_density(alpha=0.5) + facet_wrap(~Group) + theme_classic() + scale_x_log10()
figure5bsupp
mean(All_Matched_cellNumber$CellNumber)
#[1] 33.87988
mean(All_UnMatched_cellNumber$CellNumber)
#[1] 8.134875
ggsave("Figure_6B.png", plot = figure5bsupp, bg = "transparent", width = 15, height = 10)
```


## another strategy
```{r checkNumbers}
all_datasets <- list(isoform_annotations_ccRCC2_tama$TamaID,
                     isoform_annotations_ccRCC3_tama$TamaID,
                     isoform_annotations_ccRCC4_tama$TamaID,
                     isoform_annotations_ccRCC5_tama$TamaID,
                     isoform_annotations_normal_tama$TamaID)
AllMatchingIDs <- Reduce(intersect, all_datasets)
```


```{r data}
# Initialize a vector to store the count of occurrences
TamaID_count <- table(unlist(all_datasets))

# Find TamaIDs present in exactly 4 datasets
MatchingIn4 <- names(TamaID_count[TamaID_count == 4]) ## 13128

# Find TamaIDs present in exactly 3 datasets
MatchingIn3 <- names(TamaID_count[TamaID_count == 3])

# Find TamaIDs present in exactly 2 datasets
MatchingIn2 <- names(TamaID_count[TamaID_count == 2])

# Find TamaIDs unique to only one dataset
UniqueToOne <- names(TamaID_count[TamaID_count == 1])
```

```{r}
isoform_annotations_all <- rbind(isoform_annotations_normal_tama,isoform_annotations_ccRCC2_tama,isoform_annotations_ccRCC3_tama, isoform_annotations_ccRCC4_tama, isoform_annotations_ccRCC5_tama)

isoform_annotations_all_1_sample <- isoform_annotations_all[isoform_annotations_all$TamaID %in% UniqueToOne, ]
isoform_annotations_all_all_sample <- isoform_annotations_all[isoform_annotations_all$TamaID %in% AllMatchingIDs, ]
isoform_annotations_all_2_sample <- isoform_annotations_all[isoform_annotations_all$TamaID %in% MatchingIn2, ]
isoform_annotations_all_3_sample <- isoform_annotations_all[isoform_annotations_all$TamaID %in% MatchingIn3, ]
isoform_annotations_all_4_sample <- isoform_annotations_all[isoform_annotations_all$TamaID %in% MatchingIn4, ]


isoform_annotations_all_1_sample_mean <- isoform_annotations_all_1_sample %>% dplyr::select(TamaID, CellNumber, Sample) %>% dplyr::distinct() %>% dplyr::group_by(TamaID) %>% dplyr::summarise(meanCell = mean(CellNumber))
isoform_annotations_all_2_sample_mean <- isoform_annotations_all_2_sample %>% dplyr::select(TamaID, CellNumber, Sample) %>% dplyr::distinct() %>% dplyr::group_by(TamaID) %>% dplyr::summarise(meanCell = mean(CellNumber))
isoform_annotations_all_3_sample_mean <- isoform_annotations_all_3_sample %>% dplyr::select(TamaID, CellNumber, Sample) %>% dplyr::distinct() %>% dplyr::group_by(TamaID) %>% dplyr::summarise(meanCell = mean(CellNumber))
isoform_annotations_all_4_sample_mean <- isoform_annotations_all_4_sample %>% dplyr::select(TamaID, CellNumber, Sample) %>% dplyr::distinct() %>% dplyr::group_by(TamaID) %>% dplyr::summarise(meanCell = mean(CellNumber))
isoform_annotations_all_5_sample_mean <- isoform_annotations_all_all_sample %>% dplyr::select(TamaID, CellNumber, Sample) %>% dplyr::distinct() %>% dplyr::group_by(TamaID) %>% dplyr::summarise(meanCell = mean(CellNumber))

isoform_annotations_all_1_sample_mean$Status <- 'Found in 1 samples'
isoform_annotations_all_2_sample_mean$Status <- 'Found in 2 samples'
isoform_annotations_all_3_sample_mean$Status <- 'Found in 3 samples'
isoform_annotations_all_4_sample_mean$Status <- 'Found in 4 samples'
isoform_annotations_all_5_sample_mean$Status <- 'Found in all samples'


All_Matched_UnMatched_newstrategy <- rbind(isoform_annotations_all_1_sample_mean, isoform_annotations_all_5_sample_mean, isoform_annotations_all_2_sample_mean, isoform_annotations_all_3_sample_mean, isoform_annotations_all_4_sample_mean)

figure5bsupp <- ggplot(All_Matched_UnMatched_newstrategy, aes(log2(meanCell), fill=Status)) + geom_density(alpha=0.5) + theme_classic() #+ scale_x_log10()
figure5bsupp
#mean(isoform_annotations_all_all_sample$meanCell)
#[1] 43.34075

#mean(isoform_annotations_all_4_sample$CellNumber)
##[1] 24.7747
#mean(isoform_annotations_all_3_sample$CellNumber)
## [1] 16.68712
#mean(isoform_annotations_all_2_sample$CellNumber)
##[1] 11.56378
#mean(isoform_annotations_all_1_sample$CellNumber)
## [1] 6.796198

library(viridis) 

library(ggpubr)
figure6_cellDist <-  ggplot(All_Matched_UnMatched_newstrategy, aes(x = factor(Status), y = meanCell, fill = Status)) +
     geom_boxplot(notch = FALSE) +
     labs(x = "Status", y = "CellNumber") + 
     #facet_grid(~ Status) +
     labs(y = "CellNumber") +
     theme_classic() + scale_y_log10() +  scale_fill_viridis_d(option = "B")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  stat_compare_means(method = "wilcox.test", aes(label = ..p.signif..), 
                     comparisons = list(c("Found in 1 samples", "Found in 2 samples"),
                                        c("Found in 2 samples", "Found in 3 samples"),
                                        c("Found in 3 samples", "Found in 4 samples"),
                                        c("Found in 4 samples", "Found in all samples")),
                     label = "p.signif") +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
ggsave("/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/Figures/Figures/Figure_4B.png", plot = figure6_cellDist, bg = "transparent", width = 10, height = 8)

All_Matched_UnMatched_newstrategy$Status <- as.factor(All_Matched_UnMatched_newstrategy$Status)

p_values <- compare_means(meanCell ~ Status, data = All_Matched_UnMatched_newstrategy, 
                          method = "wilcox.test", 
                         # group.by = "Status", 
                          comparisons = list(c("Found in 1 samples", "Found in 2 samples"),
                                             c("Found in 2 samples", "Found in 3 samples"),
                                             c("Found in 3 samples", "Found in 4 samples"),
                                             c("Found in 4 samples", "Found in all samples")))
# View the p-values
print(p_values)
```


```{r ggtranscript}
library(ggplot2)
library(ggtranscript)

annotation_gff_ccRCC2 <- read.csv('/Users/tulaykarakulak/Documents/PhD/Projects/ccRCC_Sc_lrSequencing/070923_Final/p28443_o31598_2_threecells_combined/o31598_SMRTLink_IsoSeq/p28443_o31598_2_threecells_combined/restructured.gff', sep='\t', header = FALSE)

annotation_gff_ccRCC5 <- read.csv('/Users/tulaykarakulak/Documents/PhD/Projects/ccRCC_Sc_lrSequencing/070923_Final/p28443_o31598_5/o31598_SMRTLink_IsoSeq/p28443_o31598_5/restructured.gff', sep='\t', header = FALSE)

annotation_gff_ccRCC4 <- read.csv('/Users/tulaykarakulak/Documents/PhD/Projects/ccRCC_Sc_lrSequencing/070923_Final/p28443_o31598_4/o31598_SMRTLink_IsoSeq/p28443_o31598_4/o31598_SMRTLink_IsoSeq/p28443_o31598_4/restructured.gff', sep='\t', header = FALSE)

colnames(annotation_gff_ccRCC2) <- c('seqnames', 'start', 'end', 'strand', 'type', 'gene_name','transcript_name')
colnames(annotation_gff_ccRCC5) <- c('seqnames', 'start', 'end', 'strand', 'type', 'gene_name','transcript_name')
colnames(annotation_gff_ccRCC4) <- c('seqnames', 'start', 'end', 'strand', 'type', 'gene_name','transcript_name')
```

```{r NNMT}
NNMT_annotations_ccRCC2 <- annotation_gff_ccRCC2[annotation_gff_ccRCC2$gene_name == 'PB.100830',] %>% dplyr::filter(type == 'exon')
NNMT_annotations_ccRCC5 <- annotation_gff_ccRCC5[annotation_gff_ccRCC5$gene_name == 'PB.139561',] %>% dplyr::filter(type == 'exon')
NNMT_annotations_ccRCC4 <- annotation_gff_ccRCC4[annotation_gff_ccRCC4$gene_name == 'PB.136593',] %>% dplyr::filter(type == 'exon')

NNMT_annotations_ccRCC2$Sample <- 'ccRCC2'
NNMT_annotations_ccRCC5$Sample <- 'ccRCC5'
NNMT_annotations_ccRCC4$Sample <- 'ccRCC4'

NNMT_annotations_ccRCC2_subset <- NNMT_annotations_ccRCC2 %>% dplyr::filter(transcript_name %in% c("PB.100830.44", "PB.100830.29", "PB.100830.2", "PB.100830.43")) # only .44 is the novel, the rest are FSM
NNMT_annotations_ccRCC5_subset <- NNMT_annotations_ccRCC5 %>% dplyr::filter(transcript_name %in% c("PB.139561.14", "PB.139561.17", "PB.139561.24","PB.139561.15")) # the last two are FSM
NNMT_annotations_ccRCC4_subset <- NNMT_annotations_ccRCC4 %>% dplyr::filter(transcript_name %in% c("PB.136593.5", "PB.136593.19", "PB.136593.14", "PB.136593.16")) # first three are FSM 

NNMT_annotations_subset <- rbind(NNMT_annotations_ccRCC2_subset, NNMT_annotations_ccRCC5_subset, NNMT_annotations_ccRCC4_subset)

novel_common_transcripts <- c("PB.100830.44", "PB.139561.14", "PB.136593.16")
novel_transcript <- c("PB.139561.17")

NNMT_annotations_subset <- NNMT_annotations_subset %>% mutate(color_group = ifelse(transcript_name %in% novel_common_transcripts, "Novel_Common", "FSM"))

NNMT_annotations_subset[NNMT_annotations_subset$transcript_name == novel_transcript, 'color_group'] <- 'Novel'

colors <- c("Novel_Common" = "#F59F01", "Novel" = "#B29F01", "FSM" = "#5172B4")

NNMT_annotations_subset %>%
    ggplot(aes(
        xstart = start,
        xend = end,
        y = transcript_name
    )) + geom_range(
      aes(fill = color_group)
    ) +
    geom_intron(
        data = to_intron(NNMT_annotations_subset, "transcript_name"),
        aes(strand = strand)
    ) +
    scale_fill_manual(values = colors) +
   xlim(114297930, 114298422)

```
