---
title: "Manuscript_3"
author: "Tülay Karakulak"
date: "2024-09-02"
output: html_document
---

Codes to generate Figure 3D, 3E and 3F, and Supplementary Figure 5A 

# Load Libraries
```{r loadLibraries}
library(Seurat)
library(dplyr)
library(ggplot2)
library(knitr)
library(rmarkdown)
library(RColorBrewer)
library(viridis)
set.seed(12345)
```


```{r readIsoforms}
ccRCC_Marker_Annotations <- read.table( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/ccRCC_CellType_Revised.tsv', sep='\t', header = TRUE)
head(ccRCC_Marker_Annotations)
```

```{r makeDF}
# make a dataframe with Barcodes and sample 
ccRCC_Marker_Annotations$BC <- sapply(strsplit(ccRCC_Marker_Annotations$CellBarcode, "-"), function(x) x[1])
head(ccRCC_Marker_Annotations)
```

```{r read_AnnotationFiles}
scisoseq_file_ccrcc2 <- readRDS('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC2_filtered2.RDS')

scisoseq_file_ccrcc4 <- readRDS('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC4_filtered2.RDS')

scisoseq_file_ccrcc5 <- readRDS('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC5_filtered2.RDS')

scisoseq_file_ccrcc3 <- readRDS('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC3_filtered2.RDS')

scisoseq_file_normal <- readRDS('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_normal_filtered2.RDS')
```


# analyse all cells
```{r assignCells}
scisoseq_file_ccrcc2$Status <- ifelse(scisoseq_file_ccrcc2$BC %in%  ccRCC_Marker_Annotations[ccRCC_Marker_Annotations$Sample == 'ccRCC2' & ccRCC_Marker_Annotations$CellType == 'ccRCC', 'BC'], 'ccRCC', 'non-ccRCC')

scisoseq_file_ccrcc4$Status <- ifelse(scisoseq_file_ccrcc4$BC %in%  ccRCC_Marker_Annotations[ccRCC_Marker_Annotations$Sample == 'ccRCC4' & ccRCC_Marker_Annotations$CellType == 'ccRCC', 'BC'], 'ccRCC', 'non-ccRCC')

scisoseq_file_ccrcc5$Status <- ifelse(scisoseq_file_ccrcc5$BC %in%  ccRCC_Marker_Annotations[ccRCC_Marker_Annotations$Sample == 'ccRCC5' & ccRCC_Marker_Annotations$CellType == 'ccRCC', 'BC'], 'ccRCC', 'non-ccRCC')

scisoseq_file_normal$Status <- 'non-ccRCC'
scisoseq_file_ccrcc3$Status <- 'non-ccRCC'
```

tama isoform IDs
```{r readRDSintersections}
IsoMatch_All <- read.csv('/Users/tulaykarakulak/Documents/PhD/Projects/ccRCC_Sc_lrSequencing/070923_Final/tama_merge/Isoform_New_IDs_Tama2.tsv', sep=' ', header=FALSE)
colnames(IsoMatch_All) <- c('TamaID', 'Sample', 'PBid')
head(IsoMatch_All)
```

Match Isoforms 
```{r matchTamaIDs}
scisoseq_file_ccrcc2_tama <- merge(scisoseq_file_ccrcc2, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC2',], by.x='pbid', by.y='PBid')
scisoseq_file_ccrcc3_tama <- merge(scisoseq_file_ccrcc3, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC3',], by.x='pbid', by.y='PBid')
scisoseq_file_ccrcc4_tama <- merge(scisoseq_file_ccrcc4, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC4',], by.x='pbid', by.y='PBid')
scisoseq_file_ccrcc5_tama <- merge(scisoseq_file_ccrcc5, IsoMatch_All[IsoMatch_All$Sample == 'ccRCC5',], by.x='pbid', by.y='PBid')
scisoseq_file_normal_tama <- merge(scisoseq_file_normal, IsoMatch_All[IsoMatch_All$Sample == 'Normal',], by.x='pbid', by.y='PBid')
```

### Supplementary Figure 5A
```{r ccRCCvsnonccRCCCells}
ccRCC2_Isoforms_ccRCC <- unique(scisoseq_file_ccrcc2_tama[scisoseq_file_ccrcc2_tama$Status == 'ccRCC', 'TamaID'])
ccRCC4_Isoforms_ccRCC <- unique(scisoseq_file_ccrcc4_tama[scisoseq_file_ccrcc4_tama$Status == 'ccRCC', 'TamaID'])
ccRCC5_Isoforms_ccRCC <- unique(scisoseq_file_ccrcc5_tama[scisoseq_file_ccrcc5_tama$Status == 'ccRCC', 'TamaID'])

ccRCC2_Isoforms_non_ccRCC <- unique(scisoseq_file_ccrcc2_tama[scisoseq_file_ccrcc2_tama$Status == 'non-ccRCC', 'TamaID'])
ccRCC4_Isoforms_non_ccRCC <- unique(scisoseq_file_ccrcc4_tama[scisoseq_file_ccrcc4_tama$Status == 'non-ccRCC', 'TamaID'])
ccRCC5_Isoforms_non_ccRCC <- unique(scisoseq_file_ccrcc5_tama[scisoseq_file_ccrcc5_tama$Status == 'non-ccRCC', 'TamaID'])

All_scisoseq_file_tama <- rbind(scisoseq_file_normal_tama, scisoseq_file_ccrcc5_tama, scisoseq_file_ccrcc4_tama, scisoseq_file_ccrcc2_tama, scisoseq_file_normal_tama, scisoseq_file_ccrcc3_tama)

#saveRDS(All_scisoseq_file_tama, 'All_scisoseq_file_tama.RDS')

# ccRCC Cell Specific Isoforms and these isoforms are not expressed in any other non-ccRCC Cells
ccRCC2_Specific <- setdiff(ccRCC2_Isoforms_ccRCC, All_scisoseq_file_tama[All_scisoseq_file_tama$Status == 'non-ccRCC', 'TamaID'])

ccRCC4_Specific <- setdiff(ccRCC4_Isoforms_ccRCC, All_scisoseq_file_tama[All_scisoseq_file_tama$Status == 'non-ccRCC', 'TamaID'])

ccRCC5_Specific <- setdiff(ccRCC5_Isoforms_ccRCC, All_scisoseq_file_tama[All_scisoseq_file_tama$Status == 'non-ccRCC', 'TamaID'])

ccRCC_All_samples <- list(
          ccRCC2=ccRCC2_Specific,
          ccRCC4=ccRCC4_Specific,
          ccRCC5=ccRCC5_Specific)

# Supplementary Figure 5A 
library(ggVennDiagram)
ggVennDiagram(ccRCC_All_samples) + scale_fill_gradient(low = "grey", high='#0a9396')
```

## Figure 3D
## Proportion of Isoforms across cell categories - some Isoforms might be found in both ccRCC and non-ccRCC cells
```{r isoformTypes}
## all Isoforms together
All_scisoseq_file_tama$category[All_scisoseq_file_tama$category %in% c('fusion', 'genic', 'moreJunctions', 'NA', 'antisense', 'intergenic')]  <- 'Other'

All_scisoseq_file_tama_ccRCC_nonccRCC <- All_scisoseq_file_tama[All_scisoseq_file_tama$Sample %in% c('ccRCC2', 'ccRCC4', 'ccRCC5'),] %>% dplyr::select(pbid, gene, category, Status, Sample)  %>% dplyr::distinct() %>% group_by(Sample, Status, category) %>%
    summarise(count = n()) %>%
    mutate(total = sum(count)) %>%
    mutate(proportion = count/total) %>% ungroup() 

color_palette <- viridis(5, option = "D")

cellTypesPro_seurat_pos_neg <- ggplot(All_scisoseq_file_tama_ccRCC_nonccRCC, aes(y = proportion, x = Status, fill = category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "CellTypes", x = "proportion", title = "Proportion of Status-specific Transcript Categories across ccRCC Status") +
  #scale_fill_manual(values = color_palette_inferno) + 
  theme_classic() +
  scale_fill_manual(values = color_palette) + 
  facet_wrap(~Sample, scales="free_x") # +
  #theme(legend.position="right")

cellTypesPro_seurat_pos_neg
```

ccRCC and nonccRCC Explicit Isoforms - it is added to the manuscript 
```{r isoformTypes2}
# Sample 2
nonccRCC2_Sample_Specific <- setdiff(ccRCC2_Isoforms_non_ccRCC, ccRCC2_Isoforms_ccRCC)
ccRCC2_Sample_Specific <- setdiff(ccRCC2_Isoforms_ccRCC, ccRCC2_Isoforms_non_ccRCC)

ccRCC2_Specific_Category <- All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% ccRCC2_Sample_Specific & All_scisoseq_file_tama$Sample == 'ccRCC2', ]
nonccRCC2_Specific_Category <- All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% nonccRCC2_Sample_Specific & All_scisoseq_file_tama$Sample == 'ccRCC2', ]


## Sample 4
nonccRCC4_Sample_Specific <- setdiff(ccRCC4_Isoforms_non_ccRCC, ccRCC4_Isoforms_ccRCC)
ccRCC4_Sample_Specific <- setdiff(ccRCC4_Isoforms_ccRCC, ccRCC4_Isoforms_non_ccRCC)

ccRCC4_Specific_Category <- All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% ccRCC4_Sample_Specific & All_scisoseq_file_tama$Sample == 'ccRCC4', ]
nonccRCC4_Specific_Category <- All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% nonccRCC4_Sample_Specific & All_scisoseq_file_tama$Sample == 'ccRCC4', ]

## Sample 5
nonccRCC5_Sample_Specific <- setdiff(ccRCC5_Isoforms_non_ccRCC, ccRCC5_Isoforms_ccRCC)
ccRCC5_Sample_Specific <- setdiff(ccRCC5_Isoforms_ccRCC, ccRCC5_Isoforms_non_ccRCC)

ccRCC5_Specific_Category <- All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% ccRCC5_Sample_Specific & All_scisoseq_file_tama$Sample == 'ccRCC5', ]
nonccRCC5_Specific_Category <- All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% nonccRCC5_Sample_Specific & All_scisoseq_file_tama$Sample == 'ccRCC5', ]

All_scisoseq_file_tama_specific <- rbind(ccRCC2_Specific_Category, ccRCC4_Specific_Category, ccRCC5_Specific_Category, nonccRCC2_Specific_Category, nonccRCC4_Specific_Category, nonccRCC5_Specific_Category)

All_scisoseq_file_tama_ccRCC_nonccRCC_explicit <- All_scisoseq_file_tama_specific[All_scisoseq_file_tama_specific$Sample %in% c('ccRCC2', 'ccRCC4', 'ccRCC5'),] %>% dplyr::select(pbid, gene, category, Status, Sample)  %>% dplyr::distinct() %>% group_by(Sample, Status, category) %>%
    summarise(count = n()) %>%
    mutate(total = sum(count)) %>%
    mutate(proportion = count/total) %>% ungroup() 

color_palette <- viridis(5, option = "D")

cellTypesPro_seurat_pos_neg_spec_explicit <- ggplot(All_scisoseq_file_tama_ccRCC_nonccRCC_explicit, aes(y = proportion, x = Status, fill = category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "CellTypes", x = "proportion", title = "Proportion of Status-specific Transcript Categories across ccRCC Status") +
  #scale_fill_manual(values = color_palette_inferno) + 
  theme_classic() +
  scale_fill_manual(values = color_palette) + 
  facet_wrap(~Sample, scales="free_x") # +
  #theme(legend.position="right")

#cellTypesPro_seurat_pos_neg_spec_explicit
#ggsave("Figure_Supplementary_ccRCC_nonccRCC.png", plot = cellTypesPro_seurat_pos_neg, bg = "transparent", width = 8, height = 5)
```

ccRCC2_Specific_Category %>% dplyr::filter(category %in% c('novel_in_catalog', 'novel_not_in_catalog')) %>% dplyr::distinct(BCrev, gene, TamaID, Sample) %>% dplyr::distinct() %>% dplyr::group_by(TamaID, Sample) %>% dplyr::summarise(count = n()) %>% dplyr::arrange(desc(count))

# number of ANGPTL4 isoforms
dim(ccRCC2_Specific_Category %>% dplyr::filter(category %in% c('novel_in_catalog', 'novel_not_in_catalog')) %>% dplyr::distinct(BCrev, gene, TamaID, Sample) %>% dplyr::distinct() %>% dplyr::group_by(TamaID, gene, Sample) %>% dplyr::summarise(count = n()) %>% dplyr::arrange(desc(count)) %>% dplyr::filter(gene == "ANGPTL4") %>% dplyr::filter(!TamaID %in% All_scisoseq_file_tama[All_scisoseq_file_tama$Status == 'non-ccRCC', 'TamaID']))
134   4

# number of NDUFA4L2 isoforms
dim(ccRCC5_Specific_Category %>% dplyr::filter(category %in% c('novel_in_catalog', 'novel_not_in_catalog')) %>% dplyr::distinct(BCrev, gene, TamaID, Sample) %>% dplyr::distinct() %>% dplyr::group_by(TamaID, gene, Sample) %>% dplyr::summarise(count = n()) %>% dplyr::arrange(desc(count)) %>% dplyr::filter(gene == "NDUFA4L2") %>% dplyr::filter(!TamaID %in% All_scisoseq_file_tama[All_scisoseq_file_tama$Status == 'non-ccRCC', 'TamaID']))
#  65  4


```{r saveFilesAnnotations}
isoform_annotations_ccRCC_2_filtered <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_ccRCC_2_filtered_w_CellNumber.RDS')

isoform_annotations_ccRCC_4_filtered <- readRDS( '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_ccRCC_4_filtered_w_CellNumber.RDS')

isoform_annotations_ccRCC_5_filtered <- readRDS('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotation#isoform_annotations_ccRCC_5_filtered_w_CellNumber.RDS')
```

# Figure 3E
```{r VennDiagram}
# install.packages("ggVennDiagram")
library(ggVennDiagram)

ccRCC2_Specific_Novel <- unique(All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% ccRCC2_Specific & All_scisoseq_file_tama$Sample == 'ccRCC2' & All_scisoseq_file_tama$category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'])

ccRCC4_Specific_Novel <- unique(All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% ccRCC4_Specific & All_scisoseq_file_tama$Sample == 'ccRCC4' & All_scisoseq_file_tama$category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'])

ccRCC5_Specific_Novel <- unique(All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% ccRCC5_Specific & All_scisoseq_file_tama$Sample == 'ccRCC5' & All_scisoseq_file_tama$category %in% c('novel_in_catalog', 'novel_not_in_catalog'), 'TamaID'])

ccRCC_All_samples_Novel <- list(
          ccRCC2=ccRCC2_Specific_Novel,
          ccRCC4=ccRCC4_Specific_Novel,
          ccRCC5=ccRCC5_Specific_Novel)

figure_3e <- ggVennDiagram(ccRCC_All_samples_Novel) + scale_fill_gradient(low = "grey", high='#0a9396')

figure_3e
#ggsave("Figure_3e.png", plot = figure_3e, bg = "transparent", width = 8, height = 10)
```


```{r SupplementaryTable2}
SuppTable2_ccRCC2 <- All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% ccRCC2_Specific & All_scisoseq_file_tama$Sample == 'ccRCC2' & All_scisoseq_file_tama$category %in% c('novel_in_catalog', 'novel_not_in_catalog'), ] %>% dplyr::select(pbid, transcript, gene, category, Status, Sample, BCrev)  %>% dplyr::distinct() %>% dplyr::group_by(pbid, transcript, gene, category, Status, Sample) %>% dplyr::summarise(CellNumber = n()) %>% dplyr::arrange(desc(CellNumber))

SuppTable2_ccRCC4 <- All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% ccRCC4_Specific & All_scisoseq_file_tama$Sample == 'ccRCC4' & All_scisoseq_file_tama$category %in% c('novel_in_catalog', 'novel_not_in_catalog'), ] %>% dplyr::select(pbid, transcript, gene, category, Status, Sample, BCrev)  %>% dplyr::distinct() %>% dplyr::group_by(pbid, transcript, gene, category, Status, Sample) %>% dplyr::summarise(CellNumber = n()) %>% dplyr::arrange(desc(CellNumber))

SuppTable2_ccRCC5 <- All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% ccRCC5_Specific & All_scisoseq_file_tama$Sample == 'ccRCC5' & All_scisoseq_file_tama$category %in% c('novel_in_catalog', 'novel_not_in_catalog'), ] %>% dplyr::select(pbid, transcript, gene, category, Status, Sample, BCrev)  %>% dplyr::distinct() %>% dplyr::group_by(pbid, transcript, gene, category, Status, Sample) %>% dplyr::summarise(CellNumber = n()) %>% dplyr::arrange(desc(CellNumber))

SuppTable2 <- rbind(SuppTable2_ccRCC2, SuppTable2_ccRCC4, SuppTable2_ccRCC5)

#write.table(SuppTable2, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision_#Codes_Supplementaries/SupplementaryTables/SupplementaryFile_2.tsv', quote = FALSE, row.names = #FALSE, sep = '\t')
```

## Figure 3G
```{r clusterProfiler}
intersect_all_tamaIDs <- unique(c(intersect(ccRCC_All_samples_Novel$ccRCC2, ccRCC_All_samples_Novel$ccRCC4), intersect(ccRCC_All_samples_Novel$ccRCC2, ccRCC_All_samples_Novel$ccRCC5), intersect(ccRCC_All_samples_Novel$ccRCC4, ccRCC_All_samples_Novel$ccRCC5)))

gene_Ids <- unique(All_scisoseq_file_tama[All_scisoseq_file_tama$TamaID %in% intersect_all_tamaIDs, 'gene']) # 591 genes

library(clusterProfiler)
library(msigdbr)
library(enrichplot)
m_df <- msigdbr(species = "Homo sapiens")  %>% 
  dplyr::select(gs_name, entrez_gene)
m_t2g <- msigdbr(species = "Homo sapiens", category = "C6") %>% 
  dplyr::select(gs_name, entrez_gene)

m_H <- msigdbr(species = "Homo sapiens", category = "H") %>% 
  dplyr::select(gs_name, entrez_gene) ## halmark gene set 

m_c4 <- msigdbr(species = "Homo sapiens", category = "C4") %>% 
  dplyr::select(gs_name, entrez_gene)

m_c2 <- msigdbr(species = "Homo sapiens", category = "C2") %>% 
  dplyr::select(gs_name, entrez_gene)


eg = bitr(gene_Ids, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")

em <- enricher(eg$ENTREZID, TERM2GENE=m_t2g)
em_c4 <- enricher(eg$ENTREZID, TERM2GENE=m_c4)
em_H <- enricher(eg$ENTREZID, TERM2GENE=m_H, qvalueCutoff = 0.1)
em_c2 <- enricher(eg$ENTREZID, TERM2GENE=m_c2, qvalueCutoff = 0.1)

WP <- enrichWP(eg$ENTREZID, organism = "Homo sapiens") 
#ora_mc4 <- enrichGO(eg$ENTREZID, organism = "Homo sapiens") 
wp_result <- dotplot(WP, showCategory=10)
emH_result <- dotplot(em_H, showCategory=10)
#wp_result
#ggsave("Figure3E.png", plot = emH_result, bg = "transparent", width = 8, height = 10)

# adapted the ClusterProfiler visualization
ggplot(em_H, showCategory = 10, aes(GeneRatio, fct_reorder(Description, GeneRatio))) + 
    geom_segment(aes(xend=0, yend = Description)) + 
    geom_point(aes(fill=p.adjust, size = Count), shape=21, stroke=1, color="black") +  # Black outline and original fill colors
    scale_fill_gradient(low = "#FF1493",  high = "#FFD700") +  # Keep the original color gradient for the fill
    scale_size_continuous(range=c(2, 10)) +  
    xlab("GeneRatio") + 
    theme_classic()
```




