---
title: "Isoform_Filtering"
author: "Tülay Karakulak"
date: "2024-08-13"
output: html_document
---

title: "Data Processing - Filter data/Select Isoforms based on squanti_qc output, TSS ratio and mitochondrial contaminations." 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r required_libs}
library(Seurat)
library(patchwork)
library(dplyr)
library(ggplot2)
library(knitr)
library(rmarkdown)
library(harmony)
library(Matrix)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(ggpubr)
set.seed(12345)
```

## read Seurat Objects
```{r ReadCountMatrices,  message=FALSE}
# Gene Level Analysis
# Initialize the Seurat object with the non-normalized data
# we use unfiltered gene file for the ccRCC cell selection!! 
# Cancer
ccRCC_2.data_gene <- Read10X(data.dir = "/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC2/scisoseq.seurat_info/genes_seurat/", gene.column = 2)
ccRCC_2 <- CreateSeuratObject(counts = ccRCC_2.data_gene, project = "ccRCC_2", min.cells = 3, min.features = 100)

ccRCC_3.data_gene <- Read10X(data.dir = "/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC3/scisoseq.seurat_info/genes_seurat/", gene.column = 2)
ccRCC_3 <- CreateSeuratObject(counts = ccRCC_3.data_gene, project = "ccRCC_3", min.cells = 3, min.features = 100)

ccRCC_4.data_gene <- Read10X(data.dir = "/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC4/scisoseq.seurat_info/genes_seurat/", gene.column = 2)
ccRCC_4 <- CreateSeuratObject(counts = ccRCC_4.data_gene, project = "ccRCC_4", min.cells = 3, min.features = 100)

ccRCC_5.data_gene <- Read10X(data.dir = "/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/ccRCC5/scisoseq.seurat_info/genes_seurat/", gene.column = 2)
ccRCC_5 <- CreateSeuratObject(counts = ccRCC_5.data_gene, project = "ccRCC_5", min.cells = 3, min.features = 100)

# Normal
healthy.data_gene <- Read10X(data.dir = "/Users/tulaykarakulak/Documents/GitHub/ccRCC_scLongRead/InputFiles/Normal/scisoseq.seurat_info/genes_seurat/", gene.column = 2)
healthy_gene <- CreateSeuratObject(counts = healthy.data_gene, project = "healthy", min.cells = 3, min.features = 100)

kidney.combined <- list(healthy = healthy_gene, ccRCC_2 = ccRCC_2,  ccRCC_3 = ccRCC_3,  ccRCC_4 = ccRCC_4,  ccRCC_5 = ccRCC_5)

## Isoform Level Analysis
# Cancer
ccRCC_2.data_isoform <- Read10X(data.dir = "/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_2_threecells_072024IsoformFiltering/isoforms_seurat/", gene.column = 1)
ccRCC_2_isoform <- CreateSeuratObject(counts = ccRCC_2.data_isoform, project = "ccRCC_2", min.cells = 1, min.features = 100)

ccRCC_3.data_isoform <- Read10X(data.dir = "/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_3_2_C01_072024IsoformFiltering/isoforms_seurat/", gene.column = 1)
ccRCC_3_isoform <- CreateSeuratObject(counts = ccRCC_3.data_isoform, project = "ccRCC_3", min.cells = 1, min.features = 100)

ccRCC_4.data_isoform <- Read10X(data.dir = "/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_4_3_D01_072024IsoformFiltering/isoforms_seurat/", gene.column = 1)
ccRCC_4_isoform <- CreateSeuratObject(counts = ccRCC_4.data_isoform, project = "ccRCC_4", min.cells = 1, min.features = 100)

ccRCC_5.data_isoform <- Read10X(data.dir = "/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_5_4_E01_072024IsoformFiltering/isoforms_seurat/", gene.column = 1)
ccRCC_5_isoform <- CreateSeuratObject(counts = ccRCC_5.data_isoform, project = "ccRCC_5", min.cells = 1, min.features = 100)

healthy.data_isoform <- Read10X(data.dir = "/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_1_threecells_072024IsoformFiltering/isoforms_seurat/", gene.column = 1)
healthy_isoform <- CreateSeuratObject(counts = healthy.data_isoform, project = "healthy", min.cells = 1, min.features = 100)

kidney.combined_isoform <- list(healthy = healthy_isoform, cRCC_2 = ccRCC_2_isoform,  ccRCC_3 = ccRCC_3_isoform,  ccRCC_4 = ccRCC_4_isoform,  ccRCC_5 = ccRCC_5_isoform)
```


## read TSS ratios from sqanti - short and long read files
```{r read_AnnotationFiles}
isoform_annotations_normal_tss <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/SQANTI_QC/o31598_SMRTLink_IsoSeq/SQANTI3_QC/Normal/scisoseq_transcripts.sorted.filtered_lite_classification.txt', header = TRUE, sep='\t')

# ccRCC_2
isoform_annotations_ccRCC_2_tss  <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/SQANTI_QC/o31598_SMRTLink_IsoSeq/SQANTI3_QC/ccRCC_2/scisoseq_transcripts.sorted.filtered_lite_classification.txt', header = TRUE, sep='\t')

# ccRCC_3
isoform_annotations_ccRCC_3_tss  <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/SQANTI_QC/o31598_SMRTLink_IsoSeq/SQANTI3_QC/ccRCC_3/scisoseq_transcripts.sorted.filtered_lite_classification.txt', header = TRUE, sep='\t')

isoform_annotations_ccRCC_4_tss  <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/SQANTI_QC/o31598_SMRTLink_IsoSeq/SQANTI3_QC/ccRCC_4/scisoseq_transcripts.sorted.filtered_lite_classification.txt', header = TRUE, sep='\t')

isoform_annotations_ccRCC_5_tss  <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/SQANTI_QC/o31598_SMRTLink_IsoSeq/SQANTI3_QC/ccRCC_5/scisoseq_transcripts.sorted.filtered_lite_classification.txt', header = TRUE, sep='\t')
```

### read Filtered SQANTI3 Files
```{r read_AnnotationFiles}
isoform_annotations_normal <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_1_threecells_072024IsoformFiltering/scisoseq_classification.filtered_lite_custom_classification.txt', header = TRUE, sep='\t')

# ccRCC_2
isoform_annotations_ccRCC_2 <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_2_threecells_072024IsoformFiltering/scisoseq_classification.filtered_lite_custom_classification.txt', header = TRUE, sep='\t')

# ccRCC_3
isoform_annotations_ccRCC_3 <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_3_2_C01_072024IsoformFiltering/scisoseq_classification.filtered_lite_custom_classification.txt', header = TRUE, sep='\t')

# ccRCC_4
isoform_annotations_ccRCC_4 <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_4_3_D01_072024IsoformFiltering/scisoseq_classification.filtered_lite_custom_classification.txt', header = TRUE, sep='\t')

# ccRCC_5
isoform_annotations_ccRCC_5 <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_5_4_E01_072024IsoformFiltering/scisoseq_classification.filtered_lite_custom_classification.txt', header = TRUE, sep='\t')
```

## select isoform IDs with TSS > 1
```{r read_AnnotationFiles}
isoform_annotations_normal_tss_isoforms <- isoform_annotations_normal_tss[isoform_annotations_normal_tss$ratio_TSS > 1, 'isoform']
isoform_annotations_ccRCC_2_tss_isoforms  <- isoform_annotations_ccRCC_2_tss[isoform_annotations_ccRCC_2_tss$ratio_TSS > 1, 'isoform']
isoform_annotations_ccRCC_3_tss_isoforms  <- isoform_annotations_ccRCC_3_tss[isoform_annotations_ccRCC_3_tss$ratio_TSS > 1, 'isoform']
isoform_annotations_ccRCC_4_tss_isoforms  <- isoform_annotations_ccRCC_4_tss[isoform_annotations_ccRCC_4_tss$ratio_TSS > 1, 'isoform']
isoform_annotations_ccRCC_5_tss_isoforms  <- isoform_annotations_ccRCC_5_tss[isoform_annotations_ccRCC_5_tss$ratio_TSS > 1, 'isoform']
```


## Filter based on the Mito content
### read file
```{r mitoContent}
mit_cont <- read.csv('/Users/tulaykarakulak/Downloads/mito content.csv', header = TRUE)
head(mit_cont)

mit_cont_cells_filtered <- mit_cont[mit_cont$pbPercentMito < 30, ]
# add Barcode IDs with '1'
#mit_cont_cells_filtered_Barcode_IDs <- mit_cont_cells_filtered %>% mutate(Barcode_ID = paste0(Barcode, '-', '1'))
```


Barcode Ids are not the same as in Seurat Object. We need to match them.
## read Seurat Files to Match
```{r read_AnnotationFiles}
scisoseq_file_normal <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_1_threecells_072024IsoformFiltering/scisoseq.annotated-prefilter.info.csv', header = TRUE, sep='\t')
# ccRCC_2
scisoseq_file_ccrcc2 <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_2_threecells_072024IsoformFiltering/scisoseq.annotated-prefilter.info.csv', header = TRUE, sep='\t')
# ccRCC_3
scisoseq_file_ccrcc3 <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_3_2_C01_072024IsoformFiltering/scisoseq.annotated-prefilter.info.csv', header = TRUE, sep='\t')

scisoseq_file_ccrcc4 <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_4_3_D01_072024IsoformFiltering/scisoseq.annotated-prefilter.info.csv', header = TRUE, sep='\t')

scisoseq_file_ccrcc5 <- read.csv('/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/o31598_SMRTLink_IsoSeq/o31598_customFilteredIsoforms/p28443_o31598_5_4_E01_072024IsoformFiltering/scisoseq.annotated-prefilter.info.csv', header = TRUE, sep='\t')
```


```{r selectCells}
Normal_Cells <- unique(paste0(scisoseq_file_normal[scisoseq_file_normal$BCrev %in% mit_cont_cells_filtered[mit_cont_cells_filtered$Sample == 'Normal', 'Barcode'], 'BC'], '-', '1'))

ccRCC2_Cells <-  unique(paste0(scisoseq_file_ccrcc2[scisoseq_file_ccrcc2$BCrev %in% mit_cont_cells_filtered[mit_cont_cells_filtered$Sample == 'ccRCC_2', 'Barcode'], 'BC'], '-', '1'))

ccRCC3_Cells <-  unique(paste0(scisoseq_file_ccrcc3[scisoseq_file_ccrcc3$BCrev %in% mit_cont_cells_filtered[mit_cont_cells_filtered$Sample == 'ccRCC_3', 'Barcode'], 'BC'], '-', '1'))

ccRCC4_Cells <-  unique(paste0(scisoseq_file_ccrcc4[scisoseq_file_ccrcc4$BCrev %in% mit_cont_cells_filtered[mit_cont_cells_filtered$Sample == 'ccRCC_4', 'Barcode'], 'BC'], '-', '1'))

ccRCC5_Cells <-  unique(paste0(scisoseq_file_ccrcc5[scisoseq_file_ccrcc5$BCrev %in% mit_cont_cells_filtered[mit_cont_cells_filtered$Sample == 'ccRCC_5', 'Barcode'], 'BC'], '-', '1'))
```


## Filter Cells and Isoforms from scisoseq_files - some isoforms might not be in the Isoform Count Matrices - ribosomal or mitochondrial isoforms are filtered
```{r FilterIsoformsandCells}
# filter cells based on mito content and filter isoforms if they are in the filtered Seurat Object
scisoseq_file_normal_filtered <- scisoseq_file_normal[scisoseq_file_normal$BCrev %in% mit_cont_cells_filtered[mit_cont_cells_filtered$Sample == 'Normal', 'Barcode'] & scisoseq_file_normal$pbid %in% intersect(isoform_annotations_normal_tss_isoforms, isoform_annotations_normal$isoform), ] # 390 cells 105974 isoforms

scisoseq_file_ccRCC2_filtered <- scisoseq_file_ccrcc2[scisoseq_file_ccrcc2$BCrev %in% mit_cont_cells_filtered[mit_cont_cells_filtered$Sample == 'ccRCC_2', 'Barcode'] & scisoseq_file_ccrcc2$pbid %in% intersect(isoform_annotations_ccRCC_2_tss_isoforms, isoform_annotations_ccRCC_2$isoform), ] # 334 cells 100707 isoforms

scisoseq_file_ccRCC3_filtered <- scisoseq_file_ccrcc3[scisoseq_file_ccrcc3$BCrev %in% mit_cont_cells_filtered[mit_cont_cells_filtered$Sample == 'ccRCC_3', 'Barcode'] & scisoseq_file_ccrcc3$pbid %in% intersect(isoform_annotations_ccRCC_3_tss_isoforms, isoform_annotations_ccRCC_3$isoform), ] # 272 cells # 59096 isoform

scisoseq_file_ccRCC4_filtered <- scisoseq_file_ccrcc4[scisoseq_file_ccrcc4$BCrev %in% mit_cont_cells_filtered[mit_cont_cells_filtered$Sample == 'ccRCC_4', 'Barcode'] & scisoseq_file_ccrcc4$pbid %in% intersect(isoform_annotations_ccRCC_4_tss_isoforms, isoform_annotations_ccRCC_4$isoform), ] # 1016 cells 86424 isoforms

scisoseq_file_ccRCC5_filtered <- scisoseq_file_ccrcc5[scisoseq_file_ccrcc5$BCrev %in% mit_cont_cells_filtered[mit_cont_cells_filtered$Sample == 'ccRCC_5', 'Barcode'] & scisoseq_file_ccrcc5$pbid %in% intersect(isoform_annotations_ccRCC_5_tss_isoforms, isoform_annotations_ccRCC_5$isoform), ] # 366 cells 94970 isoforms
```


```{r saveFiles}
saveRDS(scisoseq_file_normal_filtered, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_normal_filtered.RDS') # 101477 # 390 cells
saveRDS(scisoseq_file_ccRCC2_filtered, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC2_filtered.RDS') # 96847 # 334 cells
saveRDS(scisoseq_file_ccRCC3_filtered, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC3_filtered.RDS') # 56384 # 272 cells
saveRDS(scisoseq_file_ccRCC4_filtered, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC4_filtered.RDS') # 82847 # 1016 cells
saveRDS(scisoseq_file_ccRCC5_filtered, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC5_filtered.RDS') # 91132 # 366 cells
```


## select Isoform:Gene Names 
```{r read_AnnotationFiles}
isoform_annotations_normal_isoform_gene_tss <- scisoseq_file_normal_filtered %>% 
  dplyr::select(pbid, gene) %>% dplyr::distinct() %>%
  dplyr::mutate(isoform_gene = paste(pbid, gene, sep = ":")) %>% 
  dplyr::pull(isoform_gene) # 105974 
 
isoform_annotations_normal_isoform_gene_tss <- gsub("_", "+", isoform_annotations_normal_isoform_gene_tss)

isoform_annotations_ccRCC_2_isoform_gene_tss  <- scisoseq_file_ccRCC2_filtered %>% 
  dplyr::select(pbid, gene) %>% dplyr::distinct() %>%
  dplyr::mutate(isoform_gene = paste(pbid, gene, sep = ":")) %>% 
  dplyr::pull(isoform_gene) # 100707

isoform_annotations_ccRCC_2_isoform_gene_tss <- gsub("_", "+", isoform_annotations_ccRCC_2_isoform_gene_tss)

isoform_annotations_ccRCC_3_isoform_gene_tss  <- scisoseq_file_ccRCC3_filtered %>% 
  dplyr::select(pbid, gene) %>% dplyr::distinct() %>%
  dplyr::mutate(isoform_gene = paste(pbid, gene, sep = ":")) %>% 
  dplyr::pull(isoform_gene) # 59096

isoform_annotations_ccRCC_3_isoform_gene_tss <- gsub("_", "+", isoform_annotations_ccRCC_3_isoform_gene_tss)


isoform_annotations_ccRCC_4_isoform_gene_tss  <- scisoseq_file_ccRCC4_filtered %>% 
  dplyr::select(pbid, gene) %>% dplyr::distinct() %>%
  dplyr::mutate(isoform_gene = paste(pbid, gene, sep = ":")) %>% 
  dplyr::pull(isoform_gene) # 86424

isoform_annotations_ccRCC_4_isoform_gene_tss <- gsub("_", "+", isoform_annotations_ccRCC_4_isoform_gene_tss)


isoform_annotations_ccRCC_5_isoform_gene_tss  <- scisoseq_file_ccRCC5_filtered %>% 
  dplyr::select(pbid, gene) %>% dplyr::distinct() %>%
  dplyr::mutate(isoform_gene = paste(pbid, gene, sep = ":")) %>% 
  dplyr::pull(isoform_gene)  # 94970

isoform_annotations_ccRCC_5_isoform_gene_tss <- gsub("_", "+", isoform_annotations_ccRCC_5_isoform_gene_tss)
```


## Subset Isoforms with TSS > 1 from Seurat Obj and Cells with Mitoc. content
```{r subsetSeurat}
healthy_isoform_subset <- subset(healthy_isoform, features = isoform_annotations_normal_isoform_gene_tss, cells = Normal_Cells)  # 101942
ccRCC_2_isoform_subset <- subset(ccRCC_2_isoform, features = isoform_annotations_ccRCC_2_isoform_gene_tss, cells =  ccRCC2_Cells) # 97294
ccRCC_3_isoform_subset <- subset(ccRCC_3_isoform, features = isoform_annotations_ccRCC_3_isoform_gene_tss, cells =  ccRCC3_Cells) # 56596
ccRCC_4_isoform_subset <- subset(ccRCC_4_isoform, features = isoform_annotations_ccRCC_4_isoform_gene_tss, cells =  ccRCC4_Cells) # 83342
ccRCC_5_isoform_subset <- subset(ccRCC_5_isoform, features = isoform_annotations_ccRCC_5_isoform_gene_tss, cells =  ccRCC5_Cells) # 91734
```


```{r isolateIsoformAnnotationFiles}
isoform_annotations_normal_filtered <-  isoform_annotations_normal %>% 
  dplyr::filter(isoform %in% sapply(strsplit(rownames(healthy_isoform_subset), ":"), '[', 1)) # 101942 

isoform_annotations_ccRCC_2_filtered  <- isoform_annotations_ccRCC_2 %>% 
  dplyr::filter(isoform %in% sapply(strsplit(rownames(ccRCC_2_isoform_subset), ":"), '[', 1)) # 97294

isoform_annotations_ccRCC_3_filtered  <- isoform_annotations_ccRCC_3 %>% 
  dplyr::filter(isoform %in% sapply(strsplit(rownames(ccRCC_3_isoform_subset), ":"), '[', 1)) # 56596

isoform_annotations_ccRCC_4_filtered  <- isoform_annotations_ccRCC_4 %>% 
  dplyr::filter(isoform %in% sapply(strsplit(rownames(ccRCC_4_isoform_subset), ":"), '[', 1)) # 83342

isoform_annotations_ccRCC_5_filtered  <- isoform_annotations_ccRCC_5 %>% 
  dplyr::filter(isoform %in% sapply(strsplit(rownames(ccRCC_5_isoform_subset), ":"), '[', 1)) # 91734
```

same number of isoforms for the annotation file and scisoseq
# saved the isoforms with tss >1, filtered based on many criteria, and mito reads!
```{r saveFiles}
saveRDS(healthy_isoform_subset, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/healthy_isoform_subset.RDS')
saveRDS(ccRCC_2_isoform_subset, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/ccRCC_2_isoform_subset.RDS')
saveRDS(ccRCC_3_isoform_subset, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/ccRCC_3_isoform_subset.RDS')
saveRDS(ccRCC_4_isoform_subset, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/ccRCC_4_isoform_subset.RDS')
saveRDS(ccRCC_5_isoform_subset, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/ccRCC_5_isoform_subset.RDS')
```

```{r saveFilesAnnotations}
saveRDS(isoform_annotations_normal_filtered, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_normal_filtered.RDS')

saveRDS(isoform_annotations_ccRCC_2_filtered, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_ccRCC_2_filtered.RDS')

saveRDS(isoform_annotations_ccRCC_3_filtered, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_ccRCC_3_filtered.RDS')

saveRDS(isoform_annotations_ccRCC_4_filtered, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_ccRCC_4_filtered.RDS')

saveRDS(isoform_annotations_ccRCC_5_filtered, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/isoform_annotations_ccRCC_5_filtered.RDS')
```

## filter also scisoseq data, once more to have consistent isoforms

```{r FilterIsoformsandCells}
# filter cells based on mito content and filter isoforms if they are in the filtered Seurat Object
scisoseq_file_normal_filtered2 <- scisoseq_file_normal_filtered[scisoseq_file_normal_filtered$pbid %in% isoform_annotations_normal_filtered$isoform, ] # 390 cells 101942 isoforms

scisoseq_file_ccRCC2_filtered2 <- scisoseq_file_ccRCC2_filtered[scisoseq_file_ccRCC2_filtered$pbid %in% isoform_annotations_ccRCC_2_filtered$isoform, ] # 334 cells 97294 isoforms

scisoseq_file_ccRCC3_filtered2 <- scisoseq_file_ccRCC3_filtered[scisoseq_file_ccRCC3_filtered$pbid %in% isoform_annotations_ccRCC_3_filtered$isoform, ] # 272 cells # 56596 isoform

scisoseq_file_ccRCC4_filtered2 <- scisoseq_file_ccRCC4_filtered[scisoseq_file_ccRCC4_filtered$pbid %in% isoform_annotations_ccRCC_4_filtered$isoform, ] # 83342

scisoseq_file_ccRCC5_filtered2 <- scisoseq_file_ccRCC5_filtered[scisoseq_file_ccRCC5_filtered$pbid %in% isoform_annotations_ccRCC_5_filtered$isoform, ] # 91734
```



```{r saveFiles}
saveRDS(scisoseq_file_normal_filtered2, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_normal_filtered2.RDS') # 101477 # 390 cells
saveRDS(scisoseq_file_ccRCC2_filtered2, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC2_filtered2.RDS') # 96847 # 334 cells
saveRDS(scisoseq_file_ccRCC3_filtered2, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC3_filtered2.RDS') # 56384 # 272 cells
saveRDS(scisoseq_file_ccRCC4_filtered2, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC4_filtered2.RDS') # 82847 # 1016 cells
saveRDS(scisoseq_file_ccRCC5_filtered2, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/scisoseq_file_ccRCC5_filtered2.RDS') # 91132 # 366 cells
```

# subset gene matrices - subsetted based on mitochondria reads!
```{r subsetGeneMatrices}
healthy_gene_subset <- subset(healthy_gene, cells = Normal_Cells)  # 18028
ccRCC_2_gene_subset <- subset(ccRCC_2, cells =  ccRCC2_Cells) # 16759
ccRCC_3_gene_subset <- subset(ccRCC_3, cells =  ccRCC3_Cells) # 17366
ccRCC_4_gene_subset <- subset(ccRCC_4, cells =  ccRCC4_Cells) # 17741
ccRCC_5_gene_subset <- subset(ccRCC_5, cells =  ccRCC5_Cells) # 18074
```

```{r savegenes}
saveRDS(healthy_gene_subset, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/healthy_gene_subset.RDS')
saveRDS(ccRCC_2_gene_subset, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/ccRCC_2_gene_subset.RDS')
saveRDS(ccRCC_3_gene_subset, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/ccRCC_3_gene_subset.RDS')
saveRDS(ccRCC_4_gene_subset, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/ccRCC_4_gene_subset.RDS')
saveRDS(ccRCC_5_gene_subset, '/Users/tulaykarakulak/Desktop/Desktop2/ManuscriptSubmission/ccRCC_Revision/CustomFilteredIsoforms/SeuratObjs/ccRCC_5_gene_subset.RDS')
```

